### 2014-03-23 plos_template.tex
initial commit

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Title}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew D. MacManes$^{1}$, 
Author2$^{2}$, 
Author3$^{3,\ast}$
\\
\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

\section*{Introduction}

For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation.  One interesting example of adaptation lies in animals ability to survive desert conditions. Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive. \\

Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival. Indeed, mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.  Unlike most mammals, animals living in desert habitats are subjected to long periods of extreme heat and intense drought.  As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration. \textbf{The proposed research uses a novel approach integrating physiology, evolutionary genomics, and computational biology to better understand how animals survive in what appear to be unsurvivable conditions.} This integrative research program will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments. Its impact is broad, having implications for studies of human health, conservation, and climate change. This project both leverages and enhances existing ecological, behavioral, and genomic resources available to an active \textit{Peromyscus} research community, ultimately allowing for interesting studies of comparative functional biology not possible in traditional model organisms like \textit{Mus}.

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{template}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2014-03-24 Pero_Transcriptome.tex
add analysis files and figs

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
\documentclass[11pt]{article}
%\usepackage{natbib}
\usepackage{hyperref}
\usepackage{lineno}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Characterization of the transcriptome, nucleotide sequence polymorphism and selection in the desert adapted mouse \textit{Peromyscus eremicus}}
}
% Insert Author names, affiliations and corresponding author email.
\\
Matthew D. MacManes$^{1}$, 
Michael B. Eisen $^{2}$, 
\\
\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire Institution Name, Durham NH, USA
\\
\bf{2} HHMI and University of California, Berkeley, Berkeley, CA, USA
\\
$\ast$ E-mail: macmanes@gmail.com, @PeroMHC
\end{flushleft}

\linenumbers

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

\section*{Introduction}

For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation.  One interesting example of adaptation lies in animals ability to survive desert conditions. Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive. \\

Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival. Indeed, mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.  When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.  Unlike most mammals, animals living in desert habitats are subjected to long periods of extreme heat and intense drought.  As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration. \\

One such desert-adapted rodent, a cricetid rodent endemic to the Southwest United states is a novel model for the study of adaptation to desert environments. They have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water. These rodents have distinct advantage over other desert animals (e.g. \textit{Dipodomys}) in that they breed readily in captivity, which enables laboratory studies of the phenotype of interest. In addition the focal species is positioned in a clade of well known animals (e.g. \textit{P. californicus}, \textit{P. maniculatus} and \textit{P. polionotus}) \cite{Bradley:2007jo}. There are growing genetic and genomic resources available \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}. \\

While the elucidation of the mechanisms underlying adaptation to desert survival is beyond the scope of this manuscript, we aim here to lay the groundwork by characterizing the transcriptome from four distinct tissues (brain, liver, kidney, testes). These data will be included in current larger effort aimed at sequencing the entire genome. Further, via sequencing the renal tissue of a total of 15 additional animals, we characterize nucleotide polymorphism and genome wide patterns of natural selection. Together, these investigations will aid in our overarching goal -- to understand the genetic bases of adaptation in \textit{P. eremicus}.

\section*{Materials and Methods}

\subsection*{Animal Collection and Study Design}

Animals

\subsection*{RNA extraction and Sequencing}

Total RNA was extracted from each tissue using a TRIzol extraction (Invitrogen) following the manufacturers instructions. Because preparation of an RNA
library suitable for sequencing is dependent on having high
quality, intact RNA, a small aliquot of each total RNA extract was
analyzed on a Bioanalyzer 2100 (Agilent, Palo Alto, CA, USA).
Following confirmation of sample quality, the reference sequencing libraries
were made using the TruSeq stranded RNA prep kit
(Illumina), while an unstranded TruSeq kit was used to construct the other sequencing libraries. A unique index was ligated to each sample to allow for multiplexed
sequencing. Reference libraries (n=4 tissue types) were then pooled to
contain equimolar quantities of each individual library and
submitted for Illumina sequencing using two lanes of 150nt paired end sequencing using the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire). The remaining 14 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end across two lanes of an Illumina HiSeq 2000 at the Vincent G. Coates Gnome Center (University of California, Berkeley).

\subsection*{Sequence Data Preprocessing and Assembly}

Following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed from the dataset using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}. We concatenated sequence data from each reference tissue type and assembled them using the Trinity beta version released 16 March 2014 \cite{Haas:2013jq}. We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt. The assembly was conducted on the \textsc{Xsede} computer resource Blacklight. To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.77 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}. \\

\subsection*{Assembled Sequence Annotation}

From the filtered assembly, I extracted putative coding sequences using Transdecoder version 16Jan2014 (\url{http://transdecoder.sourceforge.net/}). These putative protein coding sequences were annotated using default settings of the blastx algorithm \cite{Camacho:2009fc} against the SwissProt database downloaded on 1 March 2014. Because transcriptome assemblies typically contain non-coding elements (e.g. ncRNA) in addition to protein coding sequence, we annotated the entire filtered dataset using the NCBI nt dataset, downloaded on 1 March 2014. Lastly, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the Pfam database \cite{Punta:2012ko}. \\

To identify sequences unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem. We estimated expression individually for the four tissues. Interesting patterns of expression, including instances where expression was limited to a single tissue type were identified. \\

\subsection*{Population Genomics}

In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population, located in Palm Desert, California. Sequence data were mapped to the reference transcriptome using bwa-mem. The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd}, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}. \\

% Results and Discussion can be combined.
\section*{Results}

\subsection*{RNA extraction, Sequencing, Assembly}

Raw assembly 743314 and 418Mb
Filtered Assembly: 130764 149Mb
cds: 64355 and 60Mb 37960 complete 4203 3prime  16880 5prime 5312 internal

\subsection*{Subsection 2}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis"

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{formatted.bib}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}

\vspace{10mm}
\textbf{\hypertarget{Table 1}{Table 1}} \\
\begin{center}
\begin{tabular}{l|c }
\textsc{Dataset}& \textsc{Num. Raw Reads} \\
\hline 
\textsc{Peer360 Testes} & 32M PE  \\
\textsc{Peer360 Liver} & 53M PE   \\
\textsc{Peer360 Kidney} & 56M PE   \\
\textsc{Peer360 Brain}  & 23M PE \\
\textsc{Peer305}  & 19M PE  \\
\textsc{Peer308}  & 15M PE  \\
\textsc{Peer319}  & 14M PE  \\
\textsc{Peer321}  & 9M SE  \\
\textsc{Peer340}  & 16M PE  \\
\textsc{Peer352}  & 14M PE  \\
\textsc{Peer354}  & 9M SE  \\
\textsc{Peer359}  & 14M PE  \\
\textsc{Peer365}  & 16M PE \\
\textsc{Peer366}  & 16M PE  \\
\textsc{Peer368}  & 14M PE  \\
\textsc{Peer369}  & 14M PE  \\
\textsc{Peer372}  & 17M SE  \\
\textsc{Peer373}  & 23M SE  \\
\textsc{Peer380}  & 16M SE  \\
\textsc{Peer382}  & 14M SE  \\
\end{tabular}
\begin{quote}
\small{Table 1. The number of sequencing reads per sample}
\end{quote}

\end{center}

\vspace{5mm}

\end{document}

### 2014-03-24 Pero_Transcriptome.tex
Monday 1126

@@
 
-\subsection*{RNA extraction, Sequencing, Assembly}
+\subsection*{RNA extraction, Sequencing, Assembly, Mapping}
 
-Raw assembly 743314 and 418Mb
-Filtered Assembly: 130764 149Mb
-cds: 64355 and 60Mb 37960 complete 4203 3prime  16880 5prime 5312 internal
+RNA was extracted from the hypothalamus, renal medulla, testes, or liver from each individual using sterile technique.
+TRIzol extraction resulted in a large amount of high quality (RIN $\geq$ 8) total RNA, which was used as input.
+Libraries were constructed as per the standard Illumina protocol, and ere sequenced as described above.
+The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321.
+Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
+\\
+
+Transcriptome assembly was accomplished using the program Trinity.
+The raw assembly contained 743314 assembled sequenced measuring 418Mb.
+This assembly was filtered using TMP \textgreater 1 as a threshold.
+The filtered assembly contained 130764 sequences measuring 149Mb.
+from this filteres dataset, I extracted 64355 putative coding sequences (60Mb).
+Of these 64355 sequences, 37960 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (16880 sequences), 3-prime end (4203 sequences), or were internal (5312 sequencing having neither stop nor start codon).
 
@@
 % The bibtex filename
-\bibliography{formatted.bib}
+\bibliography{formatted1.bib}
 

### 2014-03-25 Pero_Transcriptome.tex
tues AM

@@
 
-Animals
+To begin to understand how genes may underlie desert adaptation, I collected ~15 individuals from a single population \textit{P.
+eremicus} over a two year time period (2012-2013).
+These individuals were captured in live traps, then euthanized using isoflurane overdose and decapitation.
+Immediately post-mortem, the abdominal and pelvis organs were removed, cut in half (in the case of kidney), placed in RNAlater and flash frozen in liquid Nitrogen.
+Removal of the brain, with similar preservation techniques followed that.
+Time from euthanasia to removal of all organs never exceeded five minutes.
+Samples were transferred to a -80C freezer at a later date.
+These procedures were approved by the University of California Berkeley Animal Care and Use Committee and follow guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
 

### 2014-03-25 Pero_Transcriptome.tex
Tuesday PM

@@
 \textheight 21cm
-
+\linespread{1.2}
 % Bold the 'Figure #' in the caption and separate it with a period
@@
 For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation.
+Unlike more traditional approaches
+
  One interesting example of adaptation lies in animals ability to survive desert conditions.
@@
 
-Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival.
+Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival \cite{Walsberg:2000uu}.
 Indeed, mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
@@
 
-One such desert-adapted rodent, a cricetid rodent endemic to the Southwest United states is a novel model for the study of adaptation to desert environments.
+One such desert-adapted rodent, a cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, is a novel model for the study of adaptation to desert environments.
 They have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water.
@@
 These individuals were captured in live traps, then euthanized using isoflurane overdose and decapitation.
-Immediately post-mortem, the abdominal and pelvis organs were removed, cut in half (in the case of kidney), placed in RNAlater and flash frozen in liquid Nitrogen.
-Removal of the brain, with similar preservation techniques followed that.
+Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of kidney), placed in RNAlater and flash frozen in liquid Nitrogen.
+Removal of the brain, with similar preservation techniques, followed that.
 Time from euthanasia to removal of all organs never exceeded five minutes.
@@
 
-Total RNA was extracted from each tissue using a TRIzol extraction (Invitrogen) following the manufacturers instructions. Because preparation of an RNA
-library suitable for sequencing is dependent on having high
+Total RNA was extracted from each tissue using a TRIzol extraction (Invitrogen) following the manufacturers instructions.
+Because preparation of an RNA library suitable for sequencing is dependent on having high
 quality, intact RNA, a small aliquot of each total RNA extract was
@@
 were made using the TruSeq stranded RNA prep kit
-(Illumina), while an unstranded TruSeq kit was used to construct the other sequencing libraries. A unique index was ligated to each sample to allow for multiplexed
-sequencing. Reference libraries (n=4 tissue types) were then pooled to
+(Illumina), while an unstranded TruSeq kit was used to construct the other sequencing libraries.
+A unique index was ligated to each sample to allow for multiplexed sequencing.
+Reference libraries (n=4 tissue types) were then pooled to
 contain equimolar quantities of each individual library and
@@
 Following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed from the dataset using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
-We concatenated sequence data from each reference tissue type and assembled them using the Trinity beta version released 16 March 2014 \cite{Haas:2013jq}.
+We concatenated sequence data from each reference tissue type and assembled them jointly using the Trinity beta version released 16 March 2014 \cite{Haas:2013jq}.
 We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt.
 The assembly was conducted on the \textsc{Xsede} computer resource Blacklight.
-To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.77 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}.
+To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.75 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}.
 \\

### 2014-07-24 Pero_Transcriptome.tex
mystery update

@@
 {\Large
-\textbf{Characterization of the transcriptome, nucleotide sequence polymorphism and selection in the desert adapted mouse \textit{Peromyscus eremicus}}
+\textbf{Characterization of the transcriptome, nucleotide sequence polymorphism, and natural selection in the desert adapted mouse \textit{Peromyscus eremicus}}
 }
@@
 \\
-\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire Institution Name, Durham NH, USA
+\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire, Durham NH, USA
 \\
@@
 For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation.
-Unlike more traditional approaches
-
- One interesting example of adaptation lies in animals ability to survive desert conditions.
-Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive.
+Unlike more traditional approaches (e.g.
+QTL mapping), second generation sequencing allow researchers to access genome wide patterns of gene expression, nucleotide variation and ulitimately pattens of natural selection in non-model organisms lacking an extensive set of genomic tools.
+This approach has been used widely in recent years in characterizing X, Y and Z.
 \\
 
-Specifically, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival \cite{Walsberg:2000uu}.
-Indeed, mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
+One interesting example of adaptation lies in animal's ability to survive desert conditions.
+Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive \cite{Walsberg:2000uu}.
+Indeed, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival.
+Mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
 When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
@@
 As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration.
+Understanding these mechanismsm will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
 \\
 
-One such desert-adapted rodent, a cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, is a novel model for the study of adaptation to desert environments.
-They have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water.
+Genes responsible for the maintinance of water balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
+In addition to these studies, a ong standing interest in deserts adaptation has resulted in a number of studies that looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
+More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi} \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}.
+\\
+
+These studies provide a rich context for the current and future work, aimed at developing a synthetic understanding of the the genetic and genomic underpinnings of desert adaptation in rodents.
+As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes, brain), of a desert adapted cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, \textit{Peromyscus eremicus}.
+These animals have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water.
 These rodents have distinct advantage over other desert animals (e.g.
-\textit{Dipodomys}) in that they breed readily in captivity, which enables laboratory studies of the phenotype of interest.
+\textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest.
 In addition the focal species is positioned in a clade of well known animals (e.g.
@@
 maniculatus} and \textit{P.
-polionotus}) \cite{Bradley:2007jo}.
-There are growing genetic and genomic resources available \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}.
+polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib} which together suggest that future comparative studies are possible.
 \\

### 2014-07-27 Pero_Transcriptome.tex
sunday AM

@@
 \\
-\bf{1} Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire, Durham NH, USA
+$^{1}$ Department of Molecular, Cellular and Biomedical Sciences, University of New Hampshire, Durham NH, USA
 \\
-\bf{2} HHMI and University of California, Berkeley, Berkeley, CA, USA
+$^{2}$ HHMI and University of California, Berkeley, Berkeley, CA, USA
 \\
@@
 
-For biologists interested in understanding the relationship between fitness, genotype, and phenotype, modern sequencing technologies provide for an unprecedented opportunity to gain a deep understanding of genome level processes that together, underlie adaptation.
-Unlike more traditional approaches (e.g.
-QTL mapping), second generation sequencing allow researchers to access genome wide patterns of gene expression, nucleotide variation and ulitimately pattens of natural selection in non-model organisms lacking an extensive set of genomic tools.
-This approach has been used widely in recent years in characterizing X, Y and Z.
-\\
-
-One interesting example of adaptation lies in animal's ability to survive desert conditions.
-Here, heat \emph{and} drought provide for powerful selective forces, testing animals' ability to osmoregulate and thus to survive \cite{Walsberg:2000uu}.
+Deserts are widely considered one of Earth's harshest environments.
+Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely ton have evolved specialised mechanisms that may enhance fitness.
+While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
 Indeed, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival.
@@
 When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
-Unlike most mammals, animals living in desert habitats are subjected to long periods of extreme heat and intense drought.
-As a result, desert animals have evolved mechanisms through which physiologic homeostasis is maintained despite severe and prolonged dehydration.
 Understanding these mechanismsm will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
@@
 
-Genes responsible for the maintinance of water balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
-In addition to these studies, a ong standing interest in deserts adaptation has resulted in a number of studies that looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
+The genes and structures responsible for the maintinance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
+These studies, many of which have been enabled by newer sequencing technologies, serve as a foundation for studies of renal genomics in non-model organisms.
+In particular, because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
 More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi} \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}.
@@
 
-Following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed from the dataset using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
-We concatenated sequence data from each reference tissue type and assembled them jointly using the Trinity beta version released 16 March 2014 \cite{Haas:2013jq}.
+The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations.
+The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
+Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
+We concatenated sequence data from each reference tissue type and assembled them jointly using Trinity version released 17 July 2014 \cite{Haas:2013jq}.
 We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt.
-The assembly was conducted on the \textsc{Xsede} computer resource Blacklight.
+The assembly was conducted on a lunix workstation with 64 cores and 512Gb RAM.
 To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.75 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}.

### 2014-07-27 sync.branch.site.sh

@@ -5,7 +5,8 @@ n=1
 while [ $n -lt $total ]; do
 	var1=$(grep lnL null-model/${n}.hits.out | awk '{print $5}')
 	var2=$(grep lnL alt-model/${n}.hits.out | awk '{print $5}')
+	var0=$(grep 'foreground w' alt-model/${n}.hits.out | awk '{print $5}')
 	var3=$(awk -v VAR1=$var1 -v VAR2=$var2 'BEGIN {print 2*(VAR1-VAR2) <0 ? -2*(VAR1-VAR2) : 2*(VAR1-VAR2)}')
-	echo -e $n.hits.out '\t' branch-site '\t' $var3 >> branch-site.txt
+	echo -e $n.hits.out '\t' branch-site '\t' $var3 '\t' $var0 >> branch-site.txt
 	let n=n+1
 done

### 2014-07-27 Pero_Transcriptome.tex
1132

@@
 To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.75 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}.
+The resulting assembly was characterized using the software package transrate (\url{https://github.com/Blahah/transrate})
+
 \\
@@
 \\
+
+\subsection*{Natural Selection}
+
+To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents.
+The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}) implemented in transrate.
+These alignments were inputted into the software codeABC version 1.6.0 \cite{Lopes:2013fd}.
 
@@
 % The bibtex filename
-\bibliography{formatted1.bib}
+\bibliography{formatted.bib}
 

### 2014-07-31 Pero_Transcriptome.tex
pushing

@@
 Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
-We concatenated sequence data from each reference tissue type and assembled them jointly using Trinity version released 17 July 2014 \cite{Haas:2013jq}.
+Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}.
 We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt.
-The assembly was conducted on a lunix workstation with 64 cores and 512Gb RAM.
-To filter the raw sequence assembly, I estimated TPM for each assembled sequence using bwa-mem version 0.75 \cite{Li:2013wn} and eXpress version 1.51 \cite{Roberts:2012dh}, removing all contigs whose expression was less than TPM=1 \cite{Haas:2013jq}.
-The resulting assembly was characterized using the software package transrate (\url{https://github.com/Blahah/transrate})
-
+The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM.
+To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
+I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P.
+eremicus} dataset that are likely biological in origin.
+I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that only identical sequences were removed.
+The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}).
 \\

### 2014-08-02 Pero_Transcriptome.tex
working on results

@@
 eremicus} dataset that are likely biological in origin.
-I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that only identical sequences were removed.
+I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
 The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}).
@@
 
-Transcriptome assembly was accomplished using the program Trinity.
-The raw assembly contained 743314 assembled sequenced measuring 418Mb.
-This assembly was filtered using TMP \textgreater 1 as a threshold.
-The filtered assembly contained 130764 sequences measuring 149Mb.
-from this filteres dataset, I extracted 64355 putative coding sequences (60Mb).
+Transcriptome assembly for each tissue type was accomplished using the program Trinity.
+The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
+This assembly was filtered using a blast procedure with resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
+Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome.
+Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
+
+I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues.
+
+After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained.
+From this filtered concatenated dataset, I extracted 64355 putative coding sequences (60Mb).
 Of these 64355 sequences, 37960 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (16880 sequences), 3-prime end (4203 sequences), or were internal (5312 sequencing having neither stop nor start codon).

### 2014-08-03 Pero_Transcriptome.tex
added venn diagram

@@
 
-From the filtered assembly, I extracted putative coding sequences using Transdecoder version 16Jan2014 (\url{http://transdecoder.sourceforge.net/}).
+From the filtered assembly, I extracted putative coding sequences using Transdecoder version 4ul2014 (\url{http://transdecoder.sourceforge.net/}).
 These putative protein coding sequences were annotated using default settings of the blastx algorithm \cite{Camacho:2009fc} against the SwissProt database downloaded on 1 March 2014.
@@
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
+\\
 
-I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues.
+I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues {\hypertarget{Figure 1}{Figure 1}}.
+
+%\vspace{5mm}
+%\textbf{\hypertarget{Figure 1}{Figure 1}} \\
+\centerline{\includegraphics[width=20.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/venn.jpeg}}
+\begin{quote}
+\small{Figure 1. The Venn Diagram.}
+\end{quote}
 

### 2014-08-03 Pero_Transcriptome.tex
update
added transcoder stats

@@
 
-I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues {\hypertarget{Figure 1}{Figure 1}}.
-
 %\vspace{5mm}
@@
 
+I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues {\hypertarget{Figure 1}{Figure 1}}.
 After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained.
-From this filtered concatenated dataset, I extracted 64355 putative coding sequences (60Mb).
-Of these 64355 sequences, 37960 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (16880 sequences), 3-prime end (4203 sequences), or were internal (5312 sequencing having neither stop nor start codon).
+From this filtered concatenated dataset, I extracted 71626 putative coding sequences (72Mb).
+Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (20239 sequences), 3-prime end (6445 sequences), or were internal (6721 sequencing having neither stop nor start codon).
 

### 2014-08-03 Pero_Transcriptome.tex
1730

@@
 While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
-Indeed, the maintenance of water balance in animals is one of the most important physiologic processes, and is critical to desert survival.
-Mammals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
-When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
-Understanding these mechanismsm will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
+Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all animals, desert inhabitants or not.
+Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
+When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur, which suggests that there is strong selection for mechanisms supporting osmoregulation.
+Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
 \\
@@
 
-To begin to understand how genes may underlie desert adaptation, I collected ~15 individuals from a single population \textit{P.
+To begin to understand how genes may underlie desert adaptation, I collected 16 individuals from a single population \textit{P.
 eremicus} over a two year time period (2012-2013).
@@
 
-From the filtered assembly, I extracted putative coding sequences using Transdecoder version 4ul2014 (\url{http://transdecoder.sourceforge.net/}).
-These putative protein coding sequences were annotated using default settings of the blastx algorithm \cite{Camacho:2009fc} against the SwissProt database downloaded on 1 March 2014.
-Because transcriptome assemblies typically contain non-coding elements (e.g.
-ncRNA) in addition to protein coding sequence, we annotated the entire filtered dataset using the NCBI nt dataset, downloaded on 1 March 2014.
-Lastly, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the Pfam database \cite{Punta:2012ko}.
-\\
+The filtered assemblies were annotated using default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014.
+Amongst other things, the Ensemble transcript identifiers were used in the analysis of gene ontology, conducted in the PANTHER package \cite{Mi:2004iv}.
+Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
+Lastly, I extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
 
-To identify sequences unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem.
-We estimated expression individually for the four tissues.
-Interesting patterns of expression, including instances where expression was limited to a single tissue type were identified.
+To identify patterns of gene expression unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem.
+We estimated expression individually for the four tissues using default settings of the software eXpress \cite{Roberts:2012dh}.
+Interesting patterns of expression, including instances where expression was limited to a single tissue type were identified and visualized.
 \\
@@
 The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
-This assembly was filtered using a blast procedure with resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
+This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
 Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome.
@@
 %\textbf{\hypertarget{Figure 1}{Figure 1}} \\
-\centerline{\includegraphics[width=20.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/venn.jpeg}}
+\centerline{\includegraphics[width=15.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/venn.jpeg}}
 \begin{quote}
@@
 
-I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues {\hypertarget{Figure 1}{Figure 1}}.
+I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues ({\hypertarget{Figure 1}{Figure 1}}).
 After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained.

### 2014-08-06 Pero_Transcriptome.tex
writing pop genomics

@@
 %\doublespacing
+\newcommand{\peer}{\textit{P. eremicus}}
 
@@
 eremicus} dataset that are likely biological in origin.
+This procedure, when a reference dataset is available, is superior to a strategy employing expression-based filtering of the raw assembly.
 I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
@@
 To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents.
-The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}) implemented in transrate.
-These alignments were inputted into the software codeABC version 1.6.0 \cite{Lopes:2013fd}.
+The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
+An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
 
@@
 
-Transcriptome assembly for each tissue type was accomplished using the program Trinity.
+Transcriptome assembly for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}.
 The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
@@
 
-\subsection*{Subsection 2}
+\subsection*{Population Genomics}
+
+As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the sortware package ANGSD.
+The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
+The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 1), with a much smaller fraction being subject to neutral or
+
+\subsection*{Natural Selection}
+
+To test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages.
+The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1).
 

### 2014-08-18 Pero_Transcriptome.tex
added figure 2
update

@@
 Deserts are widely considered one of Earth's harshest environments.
-Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely ton have evolved specialised mechanisms that may enhance fitness.
+Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely to have evolved specialised mechanisms that may enhance fitness.
 While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
@@
 
-The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers recommendations.
+The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers defaultrecommendations.
 The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
@@
 To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
-I used a relaxed blastN procedure (evalue set to 10$^-10$) to identify contigs in the \textit{P.
+I used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
 eremicus} dataset that are likely biological in origin.
-This procedure, when a reference dataset is available, is superior to a strategy employing expression-based filtering of the raw assembly.
+This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1) of the raw assembly.
 I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
@@
 Sequence data were mapped to the reference transcriptome using bwa-mem.
-The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd}, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}.
+The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}.
 \\
@@
 
-As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the sortware package ANGSD.
+As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD.
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 1), with a much smaller fraction being subject to neutral or
+The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.
+
+\centerline{\includegraphics[width=18.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/tajD.jpeg}}
+\begin{quote}
+\small{Figure 2. Tajima's D}
+\end{quote}
 

### 2014-08-22 Pero_Transcriptome.tex
major update, adding tables

@@
 \usepackage{amssymb}
-
+\usepackage{chngpage}
 % graphicx package, useful for including eps and pdf graphics
@@
 
-Deserts are widely considered one of Earth's harshest environments.
-Animals living in desert environments are forced to endure intense heat and drought, and in turn, species having evolved in these environments are likely to have evolved specialised mechanisms that may enhance fitness.
-While living in deserts likely involves a large number of adaptive phenotypes, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
-Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all animals, desert inhabitants or not.
+Deserts are widely considered one of Earth's most harsh environments.
+Animals living in desert environments are forced to endure intense heat and drought, and as a result, species having evolved in these environments are likely to posses specialized mechanisms that may enhance fitness.
+While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
+Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not.
 Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
-When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur, which suggests that there is strong selection for mechanisms supporting osmoregulation.
+When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
+This process suggests that there is strong selection for mechanisms supporting osmoregulation.
 Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
@@
 
-The genes and structures responsible for the maintinance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
+The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
 These studies, many of which have been enabled by newer sequencing technologies, serve as a foundation for studies of renal genomics in non-model organisms.
@@
 Further, via sequencing the renal tissue of a total of 15 additional animals, we characterize nucleotide polymorphism and genome wide patterns of natural selection.
-Together, these investigations will aid in our overarching goal -- to understand the genetic bases of adaptation in \textit{P.
+Together, these investigations will aid in our overarching goal -- to understand the genetic bases of adaptation to deserts in \textit{P.
 eremicus}.
@@
 submitted for Illumina sequencing using two lanes of 150nt paired end sequencing using the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
-The remaining 14 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end across two lanes of an Illumina HiSeq 2000 at the Vincent G.
+The remaining 15 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end across two lanes of an Illumina HiSeq 2000 at the Vincent G.
 Coates Gnome Center (University of California, Berkeley).
@@
 
-The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers defaultrecommendations.
+The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers default recommendations.
 The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
@@
 I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
-The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}).
+%The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}).
 \\
@@
 
-To identify patterns of gene expression unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem.
+To identify patterns of gene expression unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}.
 We estimated expression individually for the four tissues using default settings of the software eXpress \cite{Roberts:2012dh}.
@@
 Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (20239 sequences), 3-prime end (6445 sequences), or were internal (6721 sequencing having neither stop nor start codon).
+\\
+
+\begin{adjustwidth}{-.35in}{-.05in}% adjust the L and R margins by 1 inch
+\begin{tabular}{ l l l l l l l}
+Transcript ID	&	Testes	&	Liver	&	Kidney	&	Brain	&	Genbank ID	&	Gene ID	\\
+Transcript\_83842	&	2.05E+03	&	6.40E+03	&	1.03E+04	&	5.47E+03	&	DQ073446.1	&	COX2	\\
+Transcript\_126459	&	1.43E+01	&	2.22E+04	&	2.77E+01	&	6.73E+00	&	XM\_006991665.1	&	Alb	\\
+Transcript\_128937	&	4.39E+00	&	1.91E+04	&	4.74E+02	&	2.23E+00	&	XM\_007627625.1	&	Apoa2	\\
+Transcript\_81233	&	1.71E+03	&	5.23E+03	&	6.11E+03	&	3.08E+03	&	XM\_006993867.1	&	Fth1	\\
+Transcript\_94125	&	3.67E+01	&	1.08E+04	&	2.09E+03	&	2.75E+00	&	XM\_006977178.1	&	CytP450	\\
+Transcript\_119945	&	5.03E+03	&	1.15E+03	&	1.33E+03	&	3.71E+03	&	XM\_008686011.1	&	Ubb	\\
+Transcript\_5977	&	4.95E+00	&	1.01E+04	&	3.05E+02	&	3.58E+02	&	XM\_006978668.1	&	Tf	\\
+Transcript\_4057	&	2.62E+01	&	9.32E+03	&	1.34E+02	&	8.38E+01	&	XM\_006994871.1	&	Apoc1	\\
+Transcript\_112523	&	4.07E+02	&	7.36E+03	&	7.78E+02	&	9.54E+02	&	XM\_006994872.1	&	Apoe	\\
+Transcript\_98376	&	1.98E+00	&	8.66E+03	&	1.02E+00	&	2.68E+00	&	XM\_006970208.1	&	Ttr	\\ \end{tabular}
+\begin{quote}
+\small{Table 2. The 10 transcripts with the highest mean TPM (transcripts per million).}
+\end{quote}  
+\end{adjustwidth}
 
@@
 \end{quote}
+
+\begin{adjustwidth}{-.75in}{-.75in}% adjust the L and R margins by 1 inch
+\begin{tabular}{ l l l l }
+\textbf{Transcript ID} & \textbf{GenBank ID} & \textbf{Description} & \textbf{Tajima's D}\\
+Transcript\_49049 & XM\_006533884.1 & heterogeneous nuclear ribonucleoprotein H1 (Hnrnph1) & 3.26\\
+Transcript\_38378 & XM\_006522973.1 & Son DNA binding protein (Son) & 3.19\\
+Transcript\_126187 & NM\_133739.2 & transmembrane protein 123 (Tmem123) & 3.02\\
+Transcript\_70953 & XM\_006539066.1 & chloride channel Kb (Clcnkb) & 2.96 \\
+Transcript\_37736 & XM\_006997718.1 & h-2 class I histocompatibility antigen & 2.92 \\
+Transcript\_21448 & XM\_006986148.1 & zinc finger protein 624-like & 2.84\\
+Transcript\_47450 & NM\_009560.2 zinc & finger protein 60 (Zfp60) & 2.82\\
+Transcript\_122250 & XM\_006539068.1 & chloride channel Kb (Clcnkb) & 2.81\\
+Transcript\_78367 & XM\_006496814.1 & CDC42 binding protein kinase alpha (Cdc42bpa) & 2.78 \\
+Transcript\_96470 & XM\_006987129.1 & interferon-inducible GTPase 1-like & 2.77 \\
+ \end{tabular}
+\begin{quote}
+\small{Table 4. The 10 transcripts with the highest values for Tajima's D, which is suggestive of positive selection.}
+\end{quote}  
+\end{adjustwidth}
+
+\begin{adjustwidth}{-.75in}{-.75in}% adjust the L and R margins by 1 inch
+\begin{tabular}{ l l l l }
+\textbf{Transcript ID} & \textbf{GenBank ID} & \textbf{Description} & \textbf{Tajima's D} \\
+Transcript\_84359 & XM\_006991127.1 & nuclear receptor coactivator 3 (Ncoa3) & -2.82\\
+Transcript\_87121 & XM\_006970128.1 & methyl-CpG binding domain protein 2 (Mbd2) & -2.82 \\
+Transcript\_125755 & EU053203.1 & alpha globin gene cluster & -2.78\\
+Transcript\_87128 & XM\_006976644.1 & membrane-associated ring finger (March5) & -2.76 \\
+Transcript\_55468 & XM\_006978377.1 & Vpr (HIV-1) binding protein (Vprbp) & -2.75\\
+Transcript\_116042 & XM\_006980811.1 & membrane associated guanylate kinase (Magi3) & -2.75  \\
+Transcript\_18966 & XM\_006982814.1 & ubiquitin protein ligase E3 component n-recognin 5 (Ubr5) & -2.75 \\
+Transcript\_122204 & XM\_008772511.1 & zinc finger protein 612 (Zfp612) & -2.75 \\
+Transcript\_100550 & XM\_006971297.1 & bromodomain adjacent to zinc finger domain, 1B (Baz1b) & -2.74\\
+Transcript\_33267 & XM\_006975561.1 & pumilio RNA-binding family member 1 (Pum1) & -2.75\\
+ \end{tabular}
+\begin{quote}
+\small{Table 3. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying selection.}
+\end{quote}  
+\end{adjustwidth}
 
@@
 % The bibtex filename
-\bibliography{formatted.bib}
+\bibliography{/Users/macmanes/Documents/pero_transcriptome/biblio.bib}
 

### 2014-08-22 Pero_Transcriptome.tex
formatting stuff

@@
 Libraries were constructed as per the standard Illumina protocol, and ere sequenced as described above.
-The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321.
+The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}).
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
 \\
+\textbf{\hypertarget{Table 1}{Table 1}} \\
+\begin{center}
+\begin{tabular}{l|c }
+\textsc{Dataset}& \textsc{Num. Raw Reads} \\
+\hline 
+\textsc{Peer360 Testes} & 32M PE  \\
+\textsc{Peer360 Liver} & 53M PE   \\
+\textsc{Peer360 Kidney} & 56M PE   \\
+\textsc{Peer360 Brain}  & 23M PE \\
+\textsc{Peer305}  & 19M PE  \\
+\textsc{Peer308}  & 15M PE  \\
+\textsc{Peer319}  & 14M PE  \\
+\textsc{Peer321}  & 9M SE  \\
+\textsc{Peer340}  & 16M PE  \\
+\textsc{Peer352}  & 14M PE  \\
+\textsc{Peer354}  & 9M SE  \\
+\textsc{Peer359}  & 14M PE  \\
+\textsc{Peer365}  & 16M PE \\
+\textsc{Peer366}  & 16M PE  \\
+\textsc{Peer368}  & 14M PE  \\
+\textsc{Peer369}  & 14M PE  \\
+\textsc{Peer372}  & 17M SE  \\
+\textsc{Peer373}  & 23M SE  \\
+\textsc{Peer380}  & 16M SE  \\
+\textsc{Peer382}  & 14M SE  \\
+\end{tabular}
+\begin{quote}
+\small{Table 1. The number of sequencing reads per sample}
+\end{quote}
+\end{center}
 
@@
 %\vspace{5mm}
-%\textbf{\hypertarget{Figure 1}{Figure 1}} \\
+\textbf{\hypertarget{Figure 1}{Figure 1}} \\
 \centerline{\includegraphics[width=15.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/venn.jpeg}}
@@
 
-\begin{adjustwidth}{-.35in}{-.05in}% adjust the L and R margins by 1 inch
+\textbf{\hypertarget{Table 2}{Table 2}} \\
+\begin{center}
+\begin{adjustwidth}{-.05in}{-.05in}% adjust the L and R margins by 1 inch
 \begin{tabular}{ l l l l l l l}
-Transcript ID	&	Testes	&	Liver	&	Kidney	&	Brain	&	Genbank ID	&	Gene ID	\\
+\textbf{Transcript ID}	&	\textbf{Testes}	&	\textbf{Liver}	&	\textbf{Kidney}	&	\textbf{Brain}	&	\textbf{Genbank ID}	&	\textbf{Gene ID}	\\
+\hline
 Transcript\_83842	&	2.05E+03	&	6.40E+03	&	1.03E+04	&	5.47E+03	&	DQ073446.1	&	COX2	\\
@@
 Transcript\_98376	&	1.98E+00	&	8.66E+03	&	1.02E+00	&	2.68E+00	&	XM\_006970208.1	&	Ttr	\\ \end{tabular}
+\end{center}
 \begin{quote}
@@
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.
-
+The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.\\
+\vspace{10mm}
+\textbf{\hypertarget{Figure 2}{Figure 2}} \\
 \centerline{\includegraphics[width=18.0\baselineskip]{/Users/macmanes/Documents/pero_transcriptome/tajD.jpeg}}
 \begin{quote}
-\small{Figure 2. Tajima's D}
+\small{Figure 2. The distribution of Tajima's D for all putative transcripts.}
 \end{quote}  
-
+\vspace{10mm}
+\textbf{\hypertarget{Table 3}{Table 3}} \\
+\begin{center}
 \begin{adjustwidth}{-.75in}{-.75in}% adjust the L and R margins by 1 inch
@@
 \textbf{Transcript ID} & \textbf{GenBank ID} & \textbf{Description} & \textbf{Tajima's D}\\
+\hline
 Transcript\_49049 & XM\_006533884.1 & heterogeneous nuclear ribonucleoprotein H1 (Hnrnph1) & 3.26\\
@@
 \begin{quote}
-\small{Table 4. The 10 transcripts with the highest values for Tajima's D, which is suggestive of positive selection.}
+\small{Table 3. The 10 transcripts with the highest values for Tajima's D, which is suggestive of positive selection.}
 \end{quote}  
 \end{adjustwidth}
-
+\end{center}
+\vspace{10mm}
+\textbf{\hypertarget{Table 4}{Table 4}} \\
+\begin{center}
 \begin{adjustwidth}{-.75in}{-.75in}% adjust the L and R margins by 1 inch
@@
 \textbf{Transcript ID} & \textbf{GenBank ID} & \textbf{Description} & \textbf{Tajima's D} \\
+\hline
 Transcript\_84359 & XM\_006991127.1 & nuclear receptor coactivator 3 (Ncoa3) & -2.82\\
@@
 \begin{quote}
-\small{Table 3. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying selection.}
+\small{Table 4. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying selection.}
 \end{quote}  
 \end{adjustwidth}
+\end{center}
 
@@
 \vspace{10mm}
-\textbf{\hypertarget{Table 1}{Table 1}} \\
-\begin{center}
-\begin{tabular}{l|c }
-\textsc{Dataset}& \textsc{Num. Raw Reads} \\
-\hline 
-\textsc{Peer360 Testes} & 32M PE  \\
-\textsc{Peer360 Liver} & 53M PE   \\
-\textsc{Peer360 Kidney} & 56M PE   \\
-\textsc{Peer360 Brain}  & 23M PE \\
-\textsc{Peer305}  & 19M PE  \\
-\textsc{Peer308}  & 15M PE  \\
-\textsc{Peer319}  & 14M PE  \\
-\textsc{Peer321}  & 9M SE  \\
-\textsc{Peer340}  & 16M PE  \\
-\textsc{Peer352}  & 14M PE  \\
-\textsc{Peer354}  & 9M SE  \\
-\textsc{Peer359}  & 14M PE  \\
-\textsc{Peer365}  & 16M PE \\
-\textsc{Peer366}  & 16M PE  \\
-\textsc{Peer368}  & 14M PE  \\
-\textsc{Peer369}  & 14M PE  \\
-\textsc{Peer372}  & 17M SE  \\
-\textsc{Peer373}  & 23M SE  \\
-\textsc{Peer380}  & 16M SE  \\
-\textsc{Peer382}  & 14M SE  \\
-\end{tabular}
-\begin{quote}
-\small{Table 1. The number of sequencing reads per sample}
-\end{quote}
 

### 2014-08-22 Pero_Transcriptome.tex
adding into about data availability

@@
 Libraries were constructed as per the standard Illumina protocol, and ere sequenced as described above.
-The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}).
+The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available on the Short Read Archive accession XXX).
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
@@
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
+All tissue-specific assemblies to be made available on Dryad.
 \\
@@
 
-I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues ({\hypertarget{Figure 1}{Figure 1}}).
-After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained.
-From this filtered concatenated dataset, I extracted 71626 putative coding sequences (72Mb).
+I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues.
+Specifically, I constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed me to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
+In addition to this, I estimated mean TMP (transcripts per million) for all transcripts.
+{\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.
+\\
+
+After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (To be available on Genbank).
+From this filtered concatenated dataset, I extracted 71626 putative coding sequences (72Mb, to be available on Dryad).
 Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (20239 sequences), 3-prime end (6445 sequences), or were internal (6721 sequencing having neither stop nor start codon).
+The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad.
 \\

### 2014-08-23 Pero_Transcriptome.tex
almost there

@@
 An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
+Significance was evaluated via use of the likelihood ratio test.
+\\
 
 % Results and Discussion can be combined.
-\section*{Results}
+\section*{Results and Discussion}
 
@@
 TRIzol extraction resulted in a large amount of high quality (RIN $\geq$ 8) total RNA, which was used as input.
-Libraries were constructed as per the standard Illumina protocol, and ere sequenced as described above.
+Libraries were constructed as per the standard Illumina protocol, and were sequenced as described above.
 The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available on the Short Read Archive accession XXX).
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
+These reads served as input for all downstream analyses.
 \\
@@
 \hline 
-\textsc{Peer360 Testes} & 32M PE  \\
-\textsc{Peer360 Liver} & 53M PE   \\
-\textsc{Peer360 Kidney} & 56M PE   \\
-\textsc{Peer360 Brain}  & 23M PE \\
+\textsc{Peer360 Testes} & 32M PE/SS  \\
+\textsc{Peer360 Liver} & 53M PE /SS  \\
+\textsc{Peer360 Kidney} & 56M PE/SS   \\
+\textsc{Peer360 Brain}  & 23M PE/SS \\
 \textsc{Peer305}  & 19M PE  \\
@@
 \begin{quote}
-\small{Table 1. The number of sequencing reads per sample}
+\small{Table 1. The number of sequencing reads per sample. PE=paired end, SS=strand specific, SE=single end sequencing. }
 \end{quote}
@@
 Specifically, I constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed me to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
-In addition to this, I estimated mean TMP (transcripts per million) for all transcripts.
+Of the 4 tissues, the kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.
+\\
+
+In addition to this, I estimated mean TMP (number of transcripts per million) for all transcripts.
 {\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.
+Several genes in this list are present predominately in a single tissue type.
+For instance Transcript_126459, Albumin is very highly expressed in the liver, but less so in the other tissues.
+It should be noted, however, that making inference based on uncorrected values for TPM is not warranted.
+Statistical testing for differential expression was not implemented due to the fact that no replicates are available.
 \\
@@
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.\\
+The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.
+\hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least.
+While the former list of genes is likely to contain transcripts experiencing positive selection in the studies population, the latter group is likely experiencing purifying selection.
+\\
+
 \vspace{10mm}
+
 \textbf{\hypertarget{Figure 2}{Figure 2}} \\
@@
 
-\section*{Discussion}
+\section*{Conclusions}
 
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis"
+As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants.
+Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
+This study aims to characterize the transcriptome of a desert-adapted rodent species, \peer.
+Specifically, we characterize the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+This study aims to provide the foundation for future experimental research.
+\\
+
+Previous studies of desert adaption in rodents have largely implicated the aquaporins \cite{Gallardo:2005fm,Huang:2001ti} in adaptation to desert environments (but see \cite{Marra:2012ec}), though most of these studies employed a candidate gene approach and therefore are limited in scope.
+The current study, which characterizes a multi-tissue transcriptome, nucleotide polymorphism, and natural selection
 
@@
 \section*{Acknowledgments}
+
+MDM was supported by a NIH Postdoctoral fellowship, then by startup funds provided by the University of New Hampshire.
+He would like to acknowledge the support of his wife and children.
+Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.
 
@@
 
-\section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
-
-\section*{Tables}
-
-\vspace{10mm}
-
-\end{center}
-
-\vspace{5mm}
-
 \end{document}

### 2014-08-23 Pero_Transcriptome.tex
this version sent to Mike

@@
 \section*{Abstract}
+
+As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants.
+Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
+This study aims to characterize the transcriptome of a desert-adapted rodent species, \textit{Peromyscus eremicus}.
+Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+We identified a set of transcripts undergoing both purifying and balancing selection based on estimates of Tajima's D.
+In addition, we used a branch site test to identify a transcript, Slc2a9 likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents.
+\\
 
@@
 To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents.
-The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
+The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
 An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
@@
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-The distribution of the results, shown in Figure 2, suggest that the vast majority of the transcriptome is under purifying selection (D \textless 0), with a much smaller fraction being subject to neutral or positive selection.
+In brief, a negative Tajima's D, a result of lower than expected average heterozygosity, is often associated with purifying or directional selection, recent selective sweep or population bottleneck.
+In contrast, a positive value for Tajima's D represents higher that expected average heterozygosity, often associated with balancing selection.
+\\
+
+The distribution of the estimates of Tajima's D for all assembled transcripts is shown in Figure 2.
+The distribution is skewed towards negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}.
 \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least.
-While the former list of genes is likely to contain transcripts experiencing positive selection in the studies population, the latter group is likely experiencing purifying selection.
+The former list of genes is likely to contain transcripts experiencing positive selection in the studied population.
+This list includes, interestingly, genes obviously related to solute and water balance (e.g.
+Clcnkb and a transmembrane protein gene), and those related to immune function (a interferon-inducible GTPase and a Class 1 MHC gene).
+The latter group, containing the transcripts whose estimates of Tajima's D is the least are likely experiencing purifying selection.
+Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences.
 \\
@@
 
-To test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+These two transcripts were chose specifically because they, the former significantly, were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.
 The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages.
 The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1).
+This limited analysis of selection is to be followed up by an analysis of genome wide patterns of natural selection.
+\\
 
@@
 Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
-This study aims to characterize the transcriptome of a desert-adapted rodent species, \peer.
-Specifically, we characterize the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
-This study aims to provide the foundation for future experimental research.
+This study aimed to characterize the transcriptome of a desert-adapted rodent species, \peer.
+Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+We identified a set of transcripts undergoing both purifying and balancing selection based on Tajima's D.
+In addition, we used a branch site test to identify a transcript, likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents.
+This study aims to provide the foundation for future experimental research, attempting to understand the genetic underpinnings of extreme osmoregulation in  desert rodents.
 \\
-
-Previous studies of desert adaption in rodents have largely implicated the aquaporins \cite{Gallardo:2005fm,Huang:2001ti} in adaptation to desert environments (but see \cite{Marra:2012ec}), though most of these studies employed a candidate gene approach and therefore are limited in scope.
-The current study, which characterizes a multi-tissue transcriptome, nucleotide polymorphism, and natural selection
 
@@
 
-MDM was supported by a NIH Postdoctoral fellowship, then by startup funds provided by the University of New Hampshire.
-He would like to acknowledge the support of his wife and children.
-Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.
+MDM would like to acknowledge the support of his wife and children. Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.
 

### 2014-08-24 Pero_Transcriptome.tex
sunday AM

@@
 Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
-This study aims to characterize the transcriptome of a desert-adapted rodent species, \textit{Peromyscus eremicus}.
-Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+One such animal, \textit{Peromyscus eremicus}, is native to the desert regions of the southwest United States, and may live it's entire life without oral fluid intake.
+As a first step towards understanding the genetics that underlie this phenotype, we present here a characterization of the transcriptome.
+We assay four tissues (kidney, liver, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
 We identified a set of transcripts undergoing both purifying and balancing selection based on estimates of Tajima's D.
-In addition, we used a branch site test to identify a transcript, Slc2a9 likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents.
+In addition, we used the branch-site test to identify a transcript, Slc2a9 likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of related non-desert rodents.
 \\
@@
 Several genes in this list are present predominately in a single tissue type.
-For instance Transcript_126459, Albumin is very highly expressed in the liver, but less so in the other tissues.
+For instance Transcript\_126459, Albumin is very highly expressed in the liver, but less so in the other tissues.
 It should be noted, however, that making inference based on uncorrected values for TPM is not warranted.
@@
 Transcript\_112523	&	4.07E+02	&	7.36E+03	&	7.78E+02	&	9.54E+02	&	XM\_006994872.1	&	Apoe	\\
-Transcript\_98376	&	1.98E+00	&	8.66E+03	&	1.02E+00	&	2.68E+00	&	XM\_006970208.1	&	Ttr	\\ \end{tabular}
+Transcript\_98376	&	1.98E+00	&	8.66E+03	&	1.02E+00	&	2.68E+00	&	XM\_006970208.1	&	Ttr	\\
+\end{tabular}
 \end{center}
@@
 In addition, we used a branch site test to identify a transcript, likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of non-desert rodents.
-This study aims to provide the foundation for future experimental research, attempting to understand the genetic underpinnings of extreme osmoregulation in  desert rodents.
 \\
@@
 
-MDM would like to acknowledge the support of his wife and children. Lastly, this work was significantly improved by comments received based on a version of the manuscript posted on biorxiv.
+MDM would like to acknowledge the support of his wife and children.
+This work was significantly improved by comments received based on a version of the manuscript posted on biorxiv, and more generally by supporters of the Open Access and Science movements.
 

### 2014-08-25 Pero_Transcriptome.tex
revisions and typos

@@
 submitted for Illumina sequencing using two lanes of 150nt paired end sequencing using the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
-The remaining 15 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end across two lanes of an Illumina HiSeq 2000 at the Vincent G.
-Coates Gnome Center (University of California, Berkeley).
+The remaining 15 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end sequencing runs across several lanes of an Illumina HiSeq 2000 at the Vincent G.
+Coates Genome Center (University of California, Berkeley).
 
@@
 
-The raw sequence reads were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers default recommendations.
-The error corrected adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers default recommendations.
+The error corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
 Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
-Reads from each tissue were assembled using Trinity version released 17 July 2014 \cite{Haas:2013jq}.
+Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}.
 We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt.
@@
 eremicus} dataset that are likely biological in origin.
-This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1) of the raw assembly.
+This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1, e.g.
+\cite{MacManes:2012bu}) of the raw assembly.
 I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
-%The resulting assembly file was characterized using the software package transrate (\url{https://github.com/Blahah/transrate}).
+This multi-fasta file was used for all subsequent analyses including annotation and mapping.
 \\
@@
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
-These reads served as input for all downstream analyses.
+These trimmed reads served as input for all downstream analyses.
 \\
@@
 \begin{quote}
-\small{Table 1. The number of sequencing reads per sample. PE=paired end, SS=strand specific, SE=single end sequencing. }
+\small{Table 1. The number of sequencing reads per sample, which is indicated by Peer{number}. PE=paired end, SS=strand specific, SE=single end sequencing. }
 \end{quote}
@@
 The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
-This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
+This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA, and \textit{P.
+maniculatus} cDNAs which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
 Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome.
@@
 Specifically, I constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed me to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
-Of the 4 tissues, the kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.
+66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
+The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.
 \\
@@
 \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least.
-The former list of genes is likely to contain transcripts experiencing positive selection in the studied population.
+The former list of genes is likely to contain transcripts experiencing balancing selection in the studied population.
 This list includes, interestingly, genes obviously related to solute and water balance (e.g.
 Clcnkb and a transmembrane protein gene), and those related to immune function (a interferon-inducible GTPase and a Class 1 MHC gene).
-The latter group, containing the transcripts whose estimates of Tajima's D is the least are likely experiencing purifying selection.
+The latter group, containing the transcripts whose estimates of Tajima's D are the smallest are likely experiencing purifying selection.
 Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences.
@@
 \begin{quote}
-\small{Table 3. The 10 transcripts with the highest values for Tajima's D, which is suggestive of positive selection.}
+\small{Table 3. The 10 transcripts with the highest values for Tajima's D, which is suggestive of balancing selection.}
 \end{quote}  
@@
 \begin{quote}
-\small{Table 4. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying selection.}
+\small{Table 4. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying or directional selection.}
 \end{quote}  

### 2014-09-12 Pero_Transcriptome.tex
to Paige

@@
 
-As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants.
-Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
-One such animal, \textit{Peromyscus eremicus}, is native to the desert regions of the southwest United States, and may live it's entire life without oral fluid intake.
-As a first step towards understanding the genetics that underlie this phenotype, we present here a characterization of the transcriptome.
+As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for their mammalian inhabitants.
+Given that osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
+One such animal, \textit{Peromyscus eremicus}, is native to the desert regions of the southwest United States, and may live its entire life without oral fluid intake.
+As a first step towards understanding the genetics that underlie this phenotype, we present here a characterization of the \peer\: transcriptome.
 We assay four tissues (kidney, liver, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
@@
 Deserts are widely considered one of Earth's most harsh environments.
-Animals living in desert environments are forced to endure intense heat and drought, and as a result, species having evolved in these environments are likely to posses specialized mechanisms that may enhance fitness.
+Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to posses specialized mechanisms to deal with them.
 While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
@@
 When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
-This process suggests that there is strong selection for mechanisms supporting osmoregulation.
+Thus there has likely been strong selection for mechanisms supporting optimal osmoregulation in species that live where water is limiting.
 Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
@@
 
-To begin to understand how genes may underlie desert adaptation, I collected 16 individuals from a single population \textit{P.
+To begin to understand how genes may underlie desert adaptation, we collected 16 individuals from a single population \textit{P.
 eremicus} over a two year time period (2012-2013).
@@
 The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM.
-To filter the raw sequence assembly, I downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
-I used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
+To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
+We used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
 eremicus} dataset that are likely biological in origin.
@@
 \cite{MacManes:2012bu}) of the raw assembly.
-I then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
+We then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
 This multi-fasta file was used for all subsequent analyses including annotation and mapping.
@@
 Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
-Lastly, I extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
+Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
 
-To identify patterns of gene expression unique to each tissue type, I mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}.
+To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}.
 We estimated expression individually for the four tissues using default settings of the software eXpress \cite{Roberts:2012dh}.
@@
 
-I then estimated gene expression on each of these tissue-specific datasets, which allowed me to understand expression patterns in the multiple tissues.
-Specifically, I constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed me to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
+We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues.
+Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed us to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
 66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
@@
 
-In addition to this, I estimated mean TMP (number of transcripts per million) for all transcripts.
+In addition to this, we estimated mean TMP (number of transcripts per million) for all transcripts.
 {\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.
@@
 After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (To be available on Genbank).
-From this filtered concatenated dataset, I extracted 71626 putative coding sequences (72Mb, to be available on Dryad).
+From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be available on Dryad).
 Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (20239 sequences), 3-prime end (6445 sequences), or were internal (6721 sequencing having neither stop nor start codon).
@@
 Transcript\_87128 & XM\_006976644.1 & membrane-associated ring finger (March5) & -2.76 \\
-Transcript\_55468 & XM\_006978377.1 & Vpr (HIV-1) binding protein (Vprbp) & -2.75\\
+Transcript\_55468 & XM\_006978377.1 & Vpr binding protein (Vprbp) & -2.75\\
 Transcript\_116042 & XM\_006980811.1 & membrane associated guanylate kinase (Magi3) & -2.75  \\
@@
 
-To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, I implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
 These two transcripts were chose specifically because they, the former significantly, were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.

### 2014-09-14 Pero_Transcriptome.tex
paige update

@@
 
-As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for their mammalian inhabitants.
+As a direct result of intense heat and aridity, deserts are thought to be among the most harsh of environments, particularly for their mammalian inhabitants.
 Given that osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
-One such animal, \textit{Peromyscus eremicus}, is native to the desert regions of the southwest United States, and may live its entire life without oral fluid intake.
-As a first step towards understanding the genetics that underlie this phenotype, we present here a characterization of the \peer\: transcriptome.
-We assay four tissues (kidney, liver, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+One such animal, \textit{Peromyscus eremicus}, is native to the desert regions of the southwest United States and may live its entire life without oral fluid intake.
+As a first step toward understanding the genetics that underlie this phenotype, we present a characterization of the \peer\: transcriptome.
+We assay four tissues (kidney, liver, brain, testes) from a single individual and supplement this with population level renal transcriptome sequencing from 15 additional animals.
 We identified a set of transcripts undergoing both purifying and balancing selection based on estimates of Tajima's D.
-In addition, we used the branch-site test to identify a transcript, Slc2a9 likely related to desert osmoregulation, undergoing enhanced selection in \peer\: relative to a set of related non-desert rodents.
+In addition, we used the branch-site test to identify a transcript -Slc2a9, likely related to desert osmoregulation -- undergoing enhanced selection in \peer\: relative to a set of related non-desert rodents.
 \\
@@
 
-Deserts are widely considered one of Earth's most harsh environments.
+Deserts are widely considered one of the harshest environments on Earth.
 Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to posses specialized mechanisms to deal with them.
 While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
-Indeed, the maintenance of water balance in animals is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not.
+Indeed, the maintenance of water balance is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not.
 Most animals are exquisitely sensitive to changes in osmolality, with slight derangement eliciting physiologic compromise.
 When the loss of water exceeds dietary intake, dehydration - and in extreme cases, death - can occur.
-Thus there has likely been strong selection for mechanisms supporting optimal osmoregulation in species that live where water is limiting.
-Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, having implications for studies of human health, conservation, and climate change.
+Thus there has likely been strong selection for mechanisms supporting optimal osmoregulation in species that live where water is limited.
+Understanding these mechanisms will significantly enhance our understanding of the physiologic processes underlying osmoregulation in extreme environments, which will have implications for studies of human health, conservation, and climate change.
 \\
@@
 The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
-These studies, many of which have been enabled by newer sequencing technologies, serve as a foundation for studies of renal genomics in non-model organisms.
-In particular, because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
-More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi} \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}.
+These studies, many of which have been enabled by newer sequencing technologies, provide a foundation for studies of renal genomics in non-model organisms.
+Because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
+More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi}, \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}.
 \\
 
-These studies provide a rich context for the current and future work, aimed at developing a synthetic understanding of the the genetic and genomic underpinnings of desert adaptation in rodents.
-As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes, brain), of a desert adapted cricetid rodent endemic to the Southwest United States \cite{Veal:2001vp}, \textit{Peromyscus eremicus}.
-These animals have a lifespan typical of small mammals, and therefore an individual may live it's entire life without ever drinking water.
+These studies provide a rich context for current and future work aimed at developing a synthetic understanding of the genetic and genomic underpinnings of desert adaptation in rodents.
+As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States \cite{Veal:2001vp}, \textit{Peromyscus eremicus}.
+These animals have a lifespan typical of small mammals, and therefore an individual may live its entire life without ever drinking water.
 These rodents have distinct advantage over other desert animals (e.g.
 \textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest.
-In addition the focal species is positioned in a clade of well known animals (e.g.
+In addition, the focal species is positioned in a clade of well known animals (e.g.
 \textit{P.
 californicus}, \textit{P.
-maniculatus} and \textit{P.
-polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib} which together suggest that future comparative studies are possible.
+maniculatus}, and \textit{P.
+polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}.
+Together, this suggests that future comparative studies are possible.
 \\
 
-While the elucidation of the mechanisms underlying adaptation to desert survival is beyond the scope of this manuscript, we aim here to lay the groundwork by characterizing the transcriptome from four distinct tissues (brain, liver, kidney, testes).
-These data will be included in current larger effort aimed at sequencing the entire genome.
-Further, via sequencing the renal tissue of a total of 15 additional animals, we characterize nucleotide polymorphism and genome wide patterns of natural selection.
-Together, these investigations will aid in our overarching goal -- to understand the genetic bases of adaptation to deserts in \textit{P.
+While the elucidation of the mechanisms underlying adaptation to desert survival is beyond the scope of this manuscript, we aim to lay the groundwork by characterizing the transcriptome from four distinct tissues (brain, liver, kidney, testes).
+These data will be included in the current larger effort aimed at sequencing the entire genome.
+Further, via sequencing the renal tissue of a total of 15 additional animals, we characterize nucleotide polymorphism and genome-wide patterns of natural selection.
+Together, these investigations will aid in our overarching goal to understand the genetic basis of adaptation to deserts in \textit{P.
 eremicus}.
@@
 
-To begin to understand how genes may underlie desert adaptation, we collected 16 individuals from a single population \textit{P.
-eremicus} over a two year time period (2012-2013).
-These individuals were captured in live traps, then euthanized using isoflurane overdose and decapitation.
-Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of kidney), placed in RNAlater and flash frozen in liquid Nitrogen.
-Removal of the brain, with similar preservation techniques, followed that.
+To begin to understand how genes may underlie desert adaptation, we collected 16 individuals from a single population of \textit{P.
+eremicus} over a two-year time period (2012-2013).
+These individuals were captured in live traps and then euthanized using isoflurane overdose and decapitation.
+Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of the kidneys), placed in RNAlater and flash frozen in liquid Nitrogen.
+Removal of the brain, with similar preservation techniques, followed.
 Time from euthanasia to removal of all organs never exceeded five minutes.
 Samples were transferred to a -80C freezer at a later date.
-These procedures were approved by the University of California Berkeley Animal Care and Use Committee and follow guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
+These procedures were approved by the University of California Berkeley Animal Care and Use Committee and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
 
@@
 
-Total RNA was extracted from each tissue using a TRIzol extraction (Invitrogen) following the manufacturers instructions.
+Total RNA was extracted from each tissue using a TRIzol extraction (Invitrogen) following the manufacturer's instructions.
 Because preparation of an RNA library suitable for sequencing is dependent on having high
@@
 contain equimolar quantities of each individual library and
-submitted for Illumina sequencing using two lanes of 150nt paired end sequencing using the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
+submitted for Illumina sequencing using two lanes of 150nt paired end sequencing employing the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
 The remaining 15 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end sequencing runs across several lanes of an Illumina HiSeq 2000 at the Vincent G.
@@
 
-The raw sequence reads corresponding to the four tissue types were error corrected using the software bless \cite{Heo:2014cb}, using kmer=25, based on the developers default recommendations.
-The error corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
-Specifically, adapter sequence contamination was removed, and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
+The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
+Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
 Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}.
-We used flags indicating the stranded nature of sequencing reads and set maximum allowable physical distance between read pairs to 999nt.
+We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt.
 The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM.
-To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}), and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
+To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
 We used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
-eremicus} dataset that are likely biological in origin.
-This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1, e.g.
-\cite{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file, reducing redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting except that sequences were clustered based on 95\% sequence similarity.
-This multi-fasta file was used for all subsequent analyses including annotation and mapping.
+eremicus} dataset likely to be biological in origin.
+This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity.
+This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
 \\
@@
 
-The filtered assemblies were annotated using default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014.
-Amongst other things, the Ensemble transcript identifiers were used in the analysis of gene ontology, conducted in the PANTHER package \cite{Mi:2004iv}.
+The filtered assemblies were annotated using the default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014.
+Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}.
 Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
@@
 To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}.
-We estimated expression individually for the four tissues using default settings of the software eXpress \cite{Roberts:2012dh}.
-Interesting patterns of expression, including instances where expression was limited to a single tissue type were identified and visualized.
+We estimated expression for the four tissues individually using default settings of the software eXpress \cite{Roberts:2012dh}.
+Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized.
 \\
@@
 
-In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population, located in Palm Desert, California.
+In addition to the reference individual sequenced at four different tissue types, we sequenced 15 other conspecific individuals from the same population in Palm Desert, California.
 Sequence data were mapped to the reference transcriptome using bwa-mem.
@@
 
-To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a resent work on desert-adapter \textit{Dipodomys} rodents.
-The coding sequence corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
-An unrooted gene tree was constructed using the online resource Clustal-Omega, and together the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
-Significance was evaluated via use of the likelihood ratio test.
+To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapter \textit{Dipodomys} rodents.
+The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
+An unrooted gene tree was constructed using the online resource Clustal-Omega, and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
+Significance was evaluated via the use of the likelihood ratio test.
 \\
@@
 
-RNA was extracted from the hypothalamus, renal medulla, testes, or liver from each individual using sterile technique.
-TRIzol extraction resulted in a large amount of high quality (RIN $\geq$ 8) total RNA, which was used as input.
-Libraries were constructed as per the standard Illumina protocol, and were sequenced as described above.
+RNA was extracted from the hypothalamus, renal medulla, testes, and liver from each individual using sterile technique.
+TRIzol extraction resulted in a large amount of high quality (RIN $\geq$ 8) total RNA, which was then used as input.
+Libraries were constructed as per the standard Illumina protocol and sequenced as described above.
 The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available on the Short Read Archive accession XXX).
-Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of reads.
+Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of the total number of reads.
 These trimmed reads served as input for all downstream analyses.
@@
 \begin{quote}
-\small{Table 1. The number of sequencing reads per sample, which is indicated by Peer{number}. PE=paired end, SS=strand specific, SE=single end sequencing. }
+\small{Table 1. The number of sequencing reads per sample, indicated by Peer[number]. PE=paired end, SS=strand specific, SE=single end sequencing. }
 \end{quote}
@@
 Transcriptome assembly for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}.
-The raw assembly for brain, liver, testes and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
-This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA, and \textit{P.
-maniculatus} cDNAs which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
-Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of kidney-derived reads to the kidney transcriptome.
+The raw assembly for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
+This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA and \textit{P.
+maniculatus} cDNAs, which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
+Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of the brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of the liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of the testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of the kidney-derived reads to the kidney transcriptome.
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
-All tissue-specific assemblies to be made available on Dryad.
+All tissue-specific assemblies are to be made available on Dryad.
 \\
@@
 We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues.
-Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}), which allowed us to visualize the proportion genes whose expression was limited to a single tissue, and those where expression was ubiquitous.
-66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
+Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous.
+66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
 The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.
@@
 {\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.
-Several genes in this list are present predominately in a single tissue type.
+Several genes in this list are predominately  present in a single tissue type.
 For instance Transcript\_126459, Albumin is very highly expressed in the liver, but less so in the other tissues.
@@
 
-After expression estimation, the filtered assemblies were concatenated together, and after removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (To be available on Genbank).
-From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be available on Dryad).
-Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while other were either truncated at the 5-prime end (20239 sequences), 3-prime end (6445 sequences), or were internal (6721 sequencing having neither stop nor start codon).
+After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (to be made available on Genbank).
+From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be made available on Dryad).
+Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while the others were either truncated at the 5-prime end (20239 sequences), the 3-prime end (6445 sequences), or were internal (6721 sequencing with neither stop nor start codon).
 The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad.
@@
 
-As detailed above, the RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD.
+As detailed above, RNAseq data from 15 individuals were mapped to the reference transcriptome with the resulting BAM files being used as input to the software package ANGSD.
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-In brief, a negative Tajima's D, a result of lower than expected average heterozygosity, is often associated with purifying or directional selection, recent selective sweep or population bottleneck.
-In contrast, a positive value for Tajima's D represents higher that expected average heterozygosity, often associated with balancing selection.
+In brief, a negative Tajima's D - a result of lower than expected average heterozygosity - is often associated with purifying or directional selection, recent selective sweep or population bottleneck.
+In contrast, a positive value for Tajima's D represents higher than expected average heterozygosity, often associated with balancing selection.
 \\
 
-The distribution of the estimates of Tajima's D for all assembled transcripts is shown in Figure 2.
-The distribution is skewed towards negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}.
-\hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is least.
+The distribution of the estimates of Tajima's D for all of the assembled transcripts is shown in Figure 2.
+The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}.
+\hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is the greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is the least.
 The former list of genes is likely to contain transcripts experiencing balancing selection in the studied population.
 This list includes, interestingly, genes obviously related to solute and water balance (e.g.
-Clcnkb and a transmembrane protein gene), and those related to immune function (a interferon-inducible GTPase and a Class 1 MHC gene).
-The latter group, containing the transcripts whose estimates of Tajima's D are the smallest are likely experiencing purifying selection.
+Clcnkb and a transmembrane protein gene) and immune function (a interferon-inducible GTPase and a Class 1 MHC gene).
+The latter group, containing transcripts whose estimates of Tajima's D are the smallest are likely experiencing purifying selection.
 Many of these transcripts are involved in core regulatory functions where mutation may have strongly negative fitness consequences.
@@
 \begin{quote}
-\small{Table 3. The 10 transcripts with the highest values for Tajima's D, which is suggestive of balancing selection.}
+\small{Table 3. The 10 transcripts with the highest values for Tajima's D, which suggests balancing selection.}
 \end{quote}  
@@
 \begin{quote}
-\small{Table 4. The 10 transcripts with the lowest values for Tajima's D, which is suggestive of purifying or directional selection.}
+\small{Table 4. The 10 transcripts with the lowest values for Tajima's D, which suggests purifying or directional selection.}
 \end{quote}  
@@
 
-To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above, setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
-These two transcripts were chose specifically because they, the former significantly, were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.
+To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above by setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.
 The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages.
@@
 
-As a direct result of intense heat and aridity, deserts are thought to be amongst the most harsh of environments, particularly for it's mammalian inhabitants.
-Given the osmoregulation can be challenging for these animals, with failure resulting in death, strong selection should be observed on genes related to the maintenance of water and solute balance.
+As a direct result of intense heat and aridity, deserts are thought to be amongst the harshest environments, particularly for mammalian inhabitants.
+Given that osmoregulation can be challenging for these animals - with failure resulting in death - strong selection should be observed on genes related to the maintenance of water and solute balance.
 This study aimed to characterize the transcriptome of a desert-adapted rodent species, \peer.
-Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, testes) from a single individual, and supplement this with population level renal transcriptome sequencing from 15 additional animals.
+Specifically, we characterized the transcriptome of four tissue types (liver, kidney, brain, and testes) from a single individual and supplemented this with population-level renal transcriptome sequencing from 15 additional animals.
 We identified a set of transcripts undergoing both purifying and balancing selection based on Tajima's D.
@@
 
-MDM would like to acknowledge the support of his wife and children.
-This work was significantly improved by comments received based on a version of the manuscript posted on biorxiv, and more generally by supporters of the Open Access and Science movements.
+%MDM would like to acknowledge the support of his wife and children.
+This work was significantly improved by comments received on a version of the manuscript posted on biorxiv and more generally by supporters of the Open Access and Science movements.
 

### 2014-09-15 Pero_Transcriptome.tex
peerJ submitted

@@
 Samples were transferred to a -80C freezer at a later date.
-These procedures were approved by the University of California Berkeley Animal Care and Use Committee and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
+These procedures were approved by the Animal Care and Use Committee located at the University of California Berkeley (protocol number R224) and University of New Hampshire (protocol number 130902) as well as the California Department of Fish and Game (protocol SC-008135) and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
 

### 2014-09-15 Pero_Transcriptome.tex
for realz

@@
 
-The raw sequence reads corresponding to the four tissue types were error corrected using the software bless \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
 The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
@@
 To filter the raw sequence assembly, we downloaded \textit{Mus musculus} cDNA and ncRNA datasets from Ensembl (\url{ftp://ftp.ensembl.org/pub/release-75/fasta/mus_musculus/}) and the \textit{Peromyscus maniculatus} reference transcriptome from NCBI (\url{ftp://ftp.ncbi.nlm.nih.gov/genomes/Peromyscus_maniculatus_bairdii/RNA/}).
-We used a blastN procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
+We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
 eremicus} dataset likely to be biological in origin.
 This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity.
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}.
-Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
+Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
 Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
 
-To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem \cite{Li:2013wn}.
-We estimated expression for the four tissues individually using default settings of the software eXpress \cite{Roberts:2012dh}.
+To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \cite{Li:2013wn}.
+We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \cite{Roberts:2012dh}.
 Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized.
@@
 To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapter \textit{Dipodomys} rodents.
-The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE \cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
+The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
 An unrooted gene tree was constructed using the online resource Clustal-Omega, and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.

### 2014-09-15 Pero_Transcriptome.tex
CTB revisions starting

@@
 Deserts are widely considered one of the harshest environments on Earth.
-Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to posses specialized mechanisms to deal with them.
+Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to possess specialized mechanisms to deal with them.
 While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
@@
 These individuals were captured in live traps and then euthanized using isoflurane overdose and decapitation.
-Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of the kidneys), placed in RNAlater and flash frozen in liquid Nitrogen.
+Immediately post-mortem, the abdominal and pelvic organs were removed, cut in half (in the case of the kidneys), placed in RNAlater and flash frozen in liquid nitrogen.
 Removal of the brain, with similar preservation techniques, followed.
@@
 This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default setting, except that sequences were clustered based on 95\% sequence similarity.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity.
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
-All tissue-specific assemblies are to be made available on Dryad.
+All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox )\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}).
 \\
@@
 
-We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues.
+We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues (Pero.tissue.xprs, will be made available on Dryad, until then on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0})).
 Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous.
@@
 
-After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (to be made available on Genbank).
+After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (to be made available on Genbank, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}.).
 From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be made available on Dryad).
 Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while the others were either truncated at the 5-prime end (20239 sequences), the 3-prime end (6445 sequences), or were internal (6721 sequencing with neither stop nor start codon).
-The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad.
+The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}..
 \\

### 2014-09-18 Pero_Transcriptome.tex
adding SRA info
before shutdown

@@
 We identified a set of transcripts undergoing both purifying and balancing selection based on estimates of Tajima's D.
-In addition, we used the branch-site test to identify a transcript -Slc2a9, likely related to desert osmoregulation -- undergoing enhanced selection in \peer\: relative to a set of related non-desert rodents.
+In addition, we used the branch-site test to identify a transcript -- Slc2a9, likely related to desert osmoregulation -- undergoing enhanced selection in \peer\: relative to a set of related non-desert rodents.
 \\
@@
 \begin{center}
-\begin{tabular}{l|c }
-\textsc{Dataset}& \textsc{Num. Raw Reads} \\
+\begin{tabular}{l|c | r}
+\textsc{Dataset}& \textsc{Num. Raw Reads} & \textsc{SRA Accession} \\
 \hline 
-\textsc{Peer360 Testes} & 32M PE/SS  \\
-\textsc{Peer360 Liver} & 53M PE /SS  \\
-\textsc{Peer360 Kidney} & 56M PE/SS   \\
-\textsc{Peer360 Brain}  & 23M PE/SS \\
-\textsc{Peer305}  & 19M PE  \\
-\textsc{Peer308}  & 15M PE  \\
-\textsc{Peer319}  & 14M PE  \\
-\textsc{Peer321}  & 9M SE  \\
-\textsc{Peer340}  & 16M PE  \\
-\textsc{Peer352}  & 14M PE  \\
-\textsc{Peer354}  & 9M SE  \\
-\textsc{Peer359}  & 14M PE  \\
-\textsc{Peer365}  & 16M PE \\
-\textsc{Peer366}  & 16M PE  \\
-\textsc{Peer368}  & 14M PE  \\
-\textsc{Peer369}  & 14M PE  \\
-\textsc{Peer372}  & 17M SE  \\
-\textsc{Peer373}  & 23M SE  \\
-\textsc{Peer380}  & 16M SE  \\
-\textsc{Peer382}  & 14M SE  \\
+\textsc{Peer360 Testes} & 32M PE/SS  & SRR1575398 \\
+\textsc{Peer360 Liver} & 53M PE /SS & SRR1575397 \\
+\textsc{Peer360 Kidney} & 56M PE/SS  & SRR1575396 \\
+\textsc{Peer360 Brain}  & 23M PE/SS & SRR1575395\\
+\textsc{Peer305}  & 19M PE  & SRR1575434  \\
+\textsc{Peer308}  & 15M PE  & SRR1575437 \\
+\textsc{Peer319}  & 14M PE  & SRR1575439\\
+\textsc{Peer321}  & 9M SE  & SRR1575441 \\
+\textsc{Peer340}  & 16M PE  & SRR1575443 \\
+\textsc{Peer352}  & 14M PE  & SRR1575464 \\
+\textsc{Peer354}  & 9M SE  & SRR1575466 \\
+\textsc{Peer359}  & 14M PE  & SRR1575492 \\
+\textsc{Peer365}  & 16M PE & SRR1575493 \\
+\textsc{Peer366}  & 16M PE & SRR1575494 \\
+\textsc{Peer368}  & 14M PE  & SRR1575624\\
+\textsc{Peer369}  & 14M PE  & SRR1575625\\
+\textsc{Peer372}  & 17M SE  & SRR1576070\\
+\textsc{Peer373}  & 23M SE  & SRR1576071\\
+\textsc{Peer380}  & 16M SE  & SRR1576072 \\
+\textsc{Peer382}  & 14M SE  & SRR1576073 \\
 \end{tabular}
 \begin{quote}
-\small{Table 1. The number of sequencing reads per sample, indicated by Peer[number]. PE=paired end, SS=strand specific, SE=single end sequencing. }
+\small{Table 1. The number of sequencing reads per sample, whose identity is indicated by Peer[number]. PE=paired end, SS=strand specific, SE=single end sequencing. }
 \end{quote}
@@
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
-All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox )\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}).
+All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}).
 \\
@@
 \begin{quote}
-\small{Figure 1. The Venn Diagram.}
+\small{Figure 1.
+The Venn Diagram, which provides a visual representation of the overlap of expression of the four tissue types.
+The majority of transcripts (66,324) are expressed in all studied tissue types.}
 \end{quote}
@@
 \end{tabular}
+\end{adjustwidth}
 \end{center}
@@
 \end{quote}
-\end{adjustwidth}
 

### 2014-09-27 Pero_Transcriptome.tex
starting revisions

@@
 We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt.
+We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \cite{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \cite{Vijay:2012gy and MacManes unpublished observations}.
+Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.
+
 The assembly was conducted on a linux workstation with 64 cores and 512Gb RAM.
@@
 66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
-The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts that may be the result of deeper sequencing.
+The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts or isoforms.
 \\

### 2014-09-27 Pero_Transcriptome.tex
reviewer2 comments

@@
 These studies provide a rich context for current and future work aimed at developing a synthetic understanding of the genetic and genomic underpinnings of desert adaptation in rodents.
-As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States \cite{Veal:2001vp}, \textit{Peromyscus eremicus}.
-These animals have a lifespan typical of small mammals, and therefore an individual may live its entire life without ever drinking water.
-These rodents have distinct advantage over other desert animals (e.g.
+As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States, \textit{Peromyscus eremicus}.
+These animals have a lifespan typical of small mammals \cite{Veal:2001vp}, and therefore an individual may live its entire life without ever drinking water.
+Additionally, they have a distinct advantage over other desert animals (e.g.
 \textit{Dipodomys}) in that they breed readily in captivity, which enables future laboratory studies of the phenotype of interest.
@@
 
-To begin to understand how genes may underlie desert adaptation, we collected 16 individuals from a single population of \textit{P.
+To begin to understand how genes may underlie desert adaptation, we collected 16 adult individuals (9 male, 7 female) from a single population of \textit{P.
 eremicus} over a two-year time period (2012-2013).
@@
 eremicus} dataset likely to be biological in origin.
-This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if TMP \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
+This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
 We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity.
@@
 
-To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapter \textit{Dipodomys} rodents.
+To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \cite{Marra:2014de}.
 The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
@@
 
-Transcriptome assembly for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}.
-The raw assembly for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
+Transcriptome assemblies for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}.
+The raw assemblies for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
 This assembly was filtered using a blastN procedure against the \textit{Mus} cDNA and ncRNA and \textit{P.
-maniculatus} cDNAs, which resulted in a final dataset containing 68331 brain-specific transcripts, 71041 liver-specific transcripts, 67340 testes-specific transcripts, and 113050 kidney-specific transcripts.
+maniculatus} cDNAs, which resulted in a final dataset containing 68331 brain-derived transcripts, 71041 liver-derived transcripts, 67340 testes-derived transcripts, and 113050 kidney-derived transcripts.
 Mapping the error-corrected adapter/quality trimmed reads to these datasets resulted in mapping 94.98\% (87.01\% properly paired) of the brain-derived reads to the brain transcriptome, 96.07\% (88.13\% properly paired) of the liver-derived reads to the liver transcriptome, 96.81\% (85.10\% properly paired) of the testes-derived reads to the testes transcriptome, and 91.87\% (83.77\% properly paired) of the kidney-derived reads to the kidney transcriptome.
@@
 
-To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above by setting the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
+To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above using alignments produced in MACSE.
+These alignments were free from indels and internal stop codons.
+We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
 These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.

### 2014-09-27 rebuttal.md
more revisions

Pero Transcriptome Revision
--

>Reviewer Comments

>Reviewer 1 (C. Titus Brown)

>Basic reporting

1. The authors do a good job at framing their work, showing why the study is needed, the limitations and the how the work will/can lead to future research.

2. The assembly and annotation steps were well thought out. Assemblies were error corrected, quality filtered and several steps were implemented for annotation using closely related species, Pfam database and extraction of putative coding sequences. **The only thing I wonder is why didn’t the authors pool the samples when assembling.** This would not change their downstream pipeline much, however, it would help to recover low expressed transcripts. (Are there any citations for this?) **Also, I do not understand if or why the addition reads for kidney were not used for assembly.**

	*I did not conduct an assembly of the pooled samples is because I believe that tissue-specific isoforms may be reconstructed with more fidelity in this manner. I make a statement to this effect on line 103:105*

	*I did not use the 15 replicate individuals in the assembly for concerns that the added polymorphism would increase run time, hardware requirements, and would decrease assembly contiguity - all effects related to a more complex de Bruijn graph. Line 99:100*

3. The author mentioned in results line 185 “The kidney appears to [be] an outlier in the number of unique sequences, though this could […] result [from] the recovery of more lowly expressed transcripts [caused by] deeper sequencing.” Why would this not also be the case for liver, which only has 3M (5%) less sequences?

	*I have removed this statement. I do wonder why this organ produced a larger number of contigs in the assembly, but a similar number of transcripts after filtering. I admit that I do not have a good explanation, and I hope a reader will..*

4. I am trying to understand the filtering process for the assembled reads. From my understanding (Page 4, lines 103:109) sequences were filtered using Blastn, (Page 4, lines 113:120) annotated using Blastn, HMMER3 and Transdecoder. Is my understanding correct? If so, why were the assembled sequences filtered with Blastn before annotated with Blastn and HMMER3? I thought the point of HMMER3 was to retain divergent sequences not detected by blastn.

	*Yes, I filtered based on blastN to the Peromyscus maniculatus and Mus musculus transcriptomes, as well as a Mus ncRNA dataset. Given the combo of P. maniculatus being so closely related (and with a high quality annotation) and Mus being more distantly related but fantastically annotated coding regions, I am confident that I am not missing a substantial number of 'real' transcripts by employing this strategy. I think it is likely that I could have recovered more contigs using Pfam, but I would worry that the sensitivity of a HMM based search may result in the recovery of 'false' transcripts, which I am relatively intolerant of.*

5. For the natural section results, I think it would be interest to add more than two genes. Perhaps the top and bottom 10 genes from the Tajima’s D analysis.
6. It would also be nice to have the various parts of the analysis in a repository, for reviewing and open science purposes.

Overall I believe it is a good paper with interesting analysis, and cool results.

Elijah K Lowe and C. Titus Brown

Experimental design

Some more explicit details to enable replication would be welcome, as described above.

Validity of the findings

No comments

Comments for the author

No comments.

Reviewer 2

Basic reporting

Major comments in basic reporting section:

1. Citation format should match the "name, year" format described for the journal, currently it is in a different, numbered format

2. Introduction, lines 46-47: In discussing that P. eremicus does not drink water, is there a study or citation that gives their lifespan and/or drinking habits? Are the authors referring back to the species account cited in the previous section?

3. In reporting the individuals captured the authors should provide some metadata such as age (juvenile vs. adult) and sex (were there equal numbers of each sex, or more of one sex than the other)?

4. In the methods lines 86-89, the specific multiplexing and number of lanes of sequencing should be reported (how many individuals were sequenced on each lane, etc.) Perhaps this information could be included in table 1.

5. The figure legend for figure 1 needs to be more descriptive and informative.

Minor Revisions:

Line 106: The abbreviation for transcripts per million (TMP) is provided here but the full term is not stated until line 189, TMP should be defined here first.

Lines 164-168. Should assembly be plural in these two sentences? As it reads, it seems that the authors are referring to one combined assembly of all 4 reference tissues, but given the numbers and the subsequent text this meaning does not seem to be correct and it should instead be 'assemblies'.

Lines 167-168: The use of 'tissue-specific' terminology is somewhat confusing as this denotes that the transcripts are unique to the tissue but this is clearly not the meaning here given lines 183-185 and figure 1.

Why aren't gene symbols provided for each of the genes in tables 3 and 4, if you are going to report gene symbols for some of the genes why not do so for all of the genes?

Line 248 and Line 250: Were p-values truly equal to 0 and 1 or are these rounded estimates, would p<.05 or p>.05 be more appropriate? This may be a matter of personal preference.
Very Minor/Grammatical revisions:

Line 48: The beginning of the sentence should probably read "These rodents have a distinct..."). This is one of several minor grammatical changes/typos that should be addressed but I will not belabor this as it is a very minor point.

Experimental design

Major comments:

1. Can the authors provide an explanation for the choice of male reproductive tissue for the reference tissues while leaving out the female reproductive tissue? Presumably one of the other sampled individuals was a female and tissue could have been harvested, yet only the testes were included in the reference transcriptome sequencing.

2. For the sentence from line 136-140 the authors later reference a paper for this, but the citation should probably be included here as well and addressed heteromyid rodents, not just Dipodomys.

3. Lines 138-142 Did the alignments produced contain insertions/deletions or internal stop codons? If so how were these treated for the PAML analysis? The results of the branch-sites test can be sensitive to alignment errors and with a small number of comparisons the alignments can be inspected manually to ensure this does not occur.

4. Line 143 Clustal-Omega is usually used as a multiple sequence aligner, can the authors provide details on what method it uses for producing a tree and whether branch lengths were provided to PAML or estimated in PAML?

5. This may be something planned in subsequent work but could the authors have provided kidney expression data for the 15 additional individuals or tested for differences in expression between the individuals of different sex?

Minor comments:

Line 95, do the authors mean PHRED < 2 or PHRED <20?

Line 109 should this say "using default settings"?

Validity of the findings

Major comments:

1. The authors mention calculating the site frequency spectrum for their data in line 133, did anything come of this analysis?

2. Is anything known about the demographic history of this population? As the authors acknowledge, the patterns used to infer selection according to Tajima's D can also be produced from demographic events and the authors should provide any data that exists on the population history. If this data does not exist the authors should state that this and include demography as a possible explanation of their data in their conclusions about tables 3 and 4.

3. Do the authors have data on Tajima's D for either of the genes tested in the branch sites test? If so they should report these values.

Minor comments:

Line 200: Do the authors mean complete coding sequence or open reading frame rather than 'complete exons'?

Comments for the author

In this study the authors characterized the transcriptome of four separate tissues for an individual Peromyscus eremicus and conducted RNA-seq of kidney tissue of an additional 15 individuals to provide transcriptomic resources for the study of osmoregulation in this desert rodent. The study is sound in its methodology and provides a significant resource, however it does require some modifications and/or clarifications about some of the methodology and data before it could be published. In the above sections I provide suggestions and/or questions that should be addressed.

© 2014, PeerJ, Inc. PO Box 614 Corte Madera, CA 94976, USA

### 2014-09-27 Pero_Transcriptome.tex
more revisions

@@
 A unique index was ligated to each sample to allow for multiplexed sequencing.
-Reference libraries (n=4 tissue types) were then pooled to
-contain equimolar quantities of each individual library and
+Reference libraries (n=4 tissue types from Peer360, a male mouse used for generating a genome sequence - not part of the current study) were then pooled to contain equimolar quantities of each individual library and
 submitted for Illumina sequencing using two lanes of 150nt paired end sequencing employing the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
@@
 The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
-An unrooted gene tree was constructed using the online resource Clustal-Omega, and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
+An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
 Significance was evaluated via the use of the likelihood ratio test.
@@
 
-In addition to this, we estimated mean TMP (number of transcripts per million) for all transcripts.
+In addition to this, we estimated mean TPM (number of transcripts per million) for all transcripts.
 {\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.

### 2014-09-27 rebuttal.md
changed ref. style

@@
 
+--- 
 1. The authors do a good job at framing their work, showing why the study is needed, the limitations and the how the work will/can lead to future research.
@@
 maniculatus being so closely related (and with a high quality annotation) and Mus being more distantly related but fantastically annotated coding regions, I am confident that I am not missing a substantial number of 'real' transcripts by employing this strategy.
-I think it is likely that I could have recovered more contigs using Pfam, but I would worry that the sensitivity of a HMM based search may result in the recovery of 'false' transcripts, which I am relatively intolerant of.*
+I think it is likely that I could have recovered more contigs using Pfam, but I would worry that the sensitivity of a HMM based search may result in the recovery of 'false' transcripts, which I am relatively intolerant of.
+I am electing to not filter based on Pfam, understanding the implications of this.*
 
 5. For the natural section results, I think it would be interest to add more than two genes. Perhaps the top and bottom 10 genes from the Tajima’s D analysis.
+
+	*I have added a few more genes, Aqp1,2,4,9.
+There is not evidence of positive selection in the eremicus lineage for any of these genes, as expected.
+The de novo (e.g.
+NOT focused on candidate genes) exploration of positive selection on P.
+eremicus will be presented in a later work, in connection with a genome sequence*
+
 6. It would also be nice to have the various parts of the analysis in a repository, for reviewing and open science purposes.
+
+	*I will make sure that the scripts are placed in a public GitHub repo.*
 
@@
 
-Elijah K Lowe and C. Titus Brown
+> Reviewer 2
 
-Experimental design
-
-Some more explicit details to enable replication would be welcome, as described above.
-
-Validity of the findings
-
-No comments
-
-Comments for the author
-
-No comments.
-
-Reviewer 2
-
-Basic reporting
+>Basic reporting
+---
 
@@
 1. Citation format should match the "name, year" format described for the journal, currently it is in a different, numbered format
+
+	*I have fixed this*
 

### 2014-09-27 Pero_Transcriptome.tex
changed ref. style

@@
 % dvipdf plos.template
-\documentclass[11pt]{article}
-%\usepackage{natbib}
+\documentclass[12pt]{article}
+\usepackage{natbib}
 \usepackage{hyperref}
@@
 % cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
+%\usepackage{cite}
 
@@
 % Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
+\bibliographystyle{apalike}
 
@@
 Animals living in desert environments are forced to endure intense heat and drought, and as a result, species living in these environments are likely to possess specialized mechanisms to deal with them.
-While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \cite{Walsberg:2000uu}.
+While living in deserts likely involves a large number of adaptive traits, the ability to osmoregulate -- to maintain the proper water and electrolyte balance -- appears to be paramount \citep{Walsberg:2000uu}.
 Indeed, the maintenance of water balance is one of the most important physiologic processes for all organisms, whether they be desert inhabitants or not.
@@
 
-The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \cite{Tatum:2009kp}, rats \cite{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \cite{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
+The genes and structures responsible for the maintenance of water and electrolyte balance are well characterized in model organisms such as mice \citep{Tatum:2009kp}, rats \citep{Romero:2007fs,Rojek:2006dg,NIELSEN:1995uq}, and humans \citep{Mobasheri:2007tt,Bedford:2003wa,NIELSEN:1999vg}.
 These studies, many of which have been enabled by newer sequencing technologies, provide a foundation for studies of renal genomics in non-model organisms.
-Because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \cite{Gallardo:2005fm}, \textit{Psammomys obesus} \cite{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \cite{Altschuler:1979du}.
-More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi}, \cite{Marra:2014de} as well as \textit{Abrothrix olivacea} \cite{Giorello:2014iv}.
+Because researchers have long been interested in desert adaptation, a number of studies have looked at the morphology or expression of single genes in the renal tissues of desert adapted rodents \textit{Phyllotis darwini} \citep{Gallardo:2005fm}, \textit{Psammomys obesus} \citep{Kaissling:1975uu}, and \textit{Perognathus penicillatus} \citep{Altschuler:1979du}.
+More recently, full renal transcriptomes have been generated for \textit{Dipodomys spectabilis} and \textit{Chaetodipus baileyi}, \citep{Marra:2014de} as well as \textit{Abrothrix olivacea} \citep{Giorello:2014iv}.
 \\
@@
 As a first step, we have sequenced, assembled, and characterized the transcriptome (using four tissue types - liver, kidney, testes and brain) of a desert adapted cricetid rodent endemic to the southwest United States, \textit{Peromyscus eremicus}.
-These animals have a lifespan typical of small mammals \cite{Veal:2001vp}, and therefore an individual may live its entire life without ever drinking water.
+These animals have a lifespan typical of small mammals \citep{Veal:2001vp}, and therefore an individual may live its entire life without ever drinking water.
 Additionally, they have a distinct advantage over other desert animals (e.g.
@@
 maniculatus}, and \textit{P.
-polionotus}) \cite{Bradley:2007jo} with growing genetic and genomic resources \cite{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}.
+polionotus}) \citep{Bradley:2007jo} with growing genetic and genomic resources \citep{Shorter:2014kw,Panhuis:2011co,Shorter:2012ib}.
 Together, this suggests that future comparative studies are possible.
@@
 Samples were transferred to a -80C freezer at a later date.
-These procedures were approved by the Animal Care and Use Committee located at the University of California Berkeley (protocol number R224) and University of New Hampshire (protocol number 130902) as well as the California Department of Fish and Game (protocol SC-008135) and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \cite{Sikes:2011dz}.
+These procedures were approved by the Animal Care and Use Committee located at the University of California Berkeley (protocol number R224) and University of New Hampshire (protocol number 130902) as well as the California Department of Fish and Game (protocol SC-008135) and followed guidelines established by the American Society of Mammalogy for the use of wild animals in research \citep{Sikes:2011dz}.
 
@@
 
-The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \cite{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
-The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \cite{MacManes:2014io} and Mbandi \cite{Christoffels:2014bg}.
-Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \cite{Lohse:2012fg}.
-Reads from each tissue were assembled using the Trinity version released 17 July 2014 \cite{Haas:2013jq}.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
+The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \citep{MacManes:2014io} and Mbandi \citep{Christoffels:2014bg}.
+Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \citep{Lohse:2012fg}.
+Reads from each tissue were assembled using the Trinity version released 17 July 2014 \citep{Haas:2013jq}.
 We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt.
-We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \cite{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \cite{Vijay:2012gy and MacManes unpublished observations}.
+We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy and MacManes unpublished observations}.
 Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.
@@
 eremicus} dataset likely to be biological in origin.
-This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \cite{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \cite{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity.
+This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity.
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 
-The filtered assemblies were annotated using the default settings of the blastN algorithm \cite{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014.
-Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \cite{Mi:2004iv}.
-Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \cite{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \cite{Punta:2012ko}.
+The filtered assemblies were annotated using the default settings of the blastN algorithm \citep{Camacho:2009fc} against the Ensembl cDNA and ncRNA datasets described above, downloaded on 1 August 2014.
+Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \citep{Mi:2004iv}.
+Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko}.
 Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
 
-To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \cite{Li:2013wn}.
-We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \cite{Roberts:2012dh}.
+To identify patterns of gene expression unique to each tissue type, we mapped sequence reads from each tissue type to the reference assembly using bwa-mem (version cloned from Github 7/1/2014)  \citep{Li:2013wn}.
+We estimated expression for the four tissues individually using default settings of the software eXpress version 1.51 \citep{Roberts:2012dh}.
 Interesting patterns of expression, including instances where expression was limited to a single tissue type, were identified and visualized.
@@
 Sequence data were mapped to the reference transcriptome using bwa-mem.
-The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \cite{Korneliussen:2013uz}.
+The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz}.
 \\
@@
 
-To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \cite{Marra:2014de}.
-The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b\cite{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \cite{Aubry:2014en}).
-An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \cite{Yang:2011bm,Yang:2007ki}.
+To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \citep{Marra:2014de}.
+The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b \citep{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \citep{Aubry:2014en}).
+An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \citep{Yang:2011bm,Yang:2007ki}.
 Significance was evaluated via the use of the likelihood ratio test.
@@
 
-Transcriptome assemblies for each tissue type was accomplished using the program Trinity \cite{Haas:2013jq}.
+Transcriptome assemblies for each tissue type was accomplished using the program Trinity \citep{Haas:2013jq}.
 The raw assemblies for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
@@
 The distribution of the estimates of Tajima's D for all of the assembled transcripts is shown in Figure 2.
-The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \cite{Chamary:2006db}.
+The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \citep{Chamary:2006db}.
 \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is the greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is the least.
@@
 We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
-These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \cite{Marra:2014de}.
+These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.
 The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages.

### 2014-09-28 rebuttal.md
cont

@@
 
-	*I have added a few more genes, Aqp1,2,4,9.
+	*I have added a few more genes, Aqp 1,2,4,9.
 There is not evidence of positive selection in the eremicus lineage for any of these genes, as expected.
@@
 >Basic reporting
+
 ---
@@
 
+	*I have changed the placement of the ref to make it more clear the issue of lifespan.
+The issue of drinking, however, seems like relatively straightforward.
+After all, these are animals that live in the desert.
+Rain may happen on rare occasions, but by definition deserts are habitats exceptionally devoid of naturally-occuring drinking water.*
+
 3. In reporting the individuals captured the authors should provide some metadata such as age (juvenile vs. adult) and sex (were there equal numbers of each sex, or more of one sex than the other)?
+
+	*This is done. Line 72*
 
@@
 
+	*This information is largely there.
+For instance, I specify that the 4 reference tissue samples were sequenced on 2 lanes of a Hiseq 2500.
+The 15 replicate samples were sequence across several lanes, so this number (how many individuals sequenced per lane) is not representative.
+Instead, the number of reads seems like the more informative number.
+*
+
 5. The figure legend for figure 1 needs to be more descriptive and informative.
+
+	*I have added more detail*
 
@@
 
-Line 106: The abbreviation for transcripts per million (TMP) is provided here but the full term is not stated until line 189, TMP should be defined here first.
+1. Line 106: The abbreviation for transcripts per million (TMP) is provided here but the full term is not stated until line 189, TMP should be defined here first.
 
+	*I have added the definition at the 1st use of the abbreviation.*
+
+2.
 Lines 164-168.
@@
 
+	*Tricky grammer issue. I have changed these terms to their plural*
+
+3.
 Lines 167-168: The use of 'tissue-specific' terminology is somewhat confusing as this denotes that the transcripts are unique to the tissue but this is clearly not the meaning here given lines 183-185 and figure 1.
 
-Why aren't gene symbols provided for each of the genes in tables 3 and 4, if you are going to report gene symbols for some of the genes why not do so for all of the genes?
+	*Excellent point! I have modified the text. Lines 198:202. It now reads brain-derived*
 
-Line 248 and Line 250: Were p-values truly equal to 0 and 1 or are these rounded estimates, would p<.05 or p>.05 be more appropriate? This may be a matter of personal preference.
-Very Minor/Grammatical revisions:
+4. Why aren't gene symbols provided for each of the genes in tables 3 and 4, if you are going to report gene symbols for some of the genes why not do so for all of the genes?
 
+	*Genes were given symbols when symbols exist. For some, e.g. h-2 class I histocompatibility antigen XM_006997718.1 has the symbol "LOC102911283". This does not seem informative, so it was left off.*
+
+5. Line 248 and Line 250: Were p-values truly equal to 0 and 1 or are these rounded estimates, would p<.05 or p>.05 be more appropriate? This may be a matter of personal preference.
+
+	*These are numbers reported by PAML, so I think it is appropriate to leave them standing as is*
+
+> Very Minor/Grammatical revisions:
+
+6.
 Line 48: The beginning of the sentence should probably read "These rodents have a distinct...").
@@
 
-Experimental design
+	*Corrected here and a few other places*
 
-Major comments:
+>Experimental design
+
+>Major comments:
 
@@
 
+	*Thanks for bring this up.
+The 4 tissues were from the male animal that has been sequenced for genome assembly - these 4 tissues will aid in the annotation process.
+From this animal, I did not sequence all available tissues (e.g.
+no lung, skeletal muscle, large intestines, etc) secondary to financial constraint.
+Had I chosen a female animal for genome sequencing, ovary would have been included.*
+
 2. For the sentence from line 136-140 the authors later reference a paper for this, but the citation should probably be included here as well and addressed heteromyid rodents, not just Dipodomys.
+
+	*Excellent point. I have added the citation here and changed Dipodomys to Hetermyid to more accurately reflect the work done in that paper.*
 
@@
 
+	*The alignments were visually inspected.
+While some indels do occur (especially in Homo relative to the other rodents), these all occur in units of 3, which to me at least suggests that they are likely accurate.
+The alignment column in which premature stop codons are treated as missing data in PAML.
+Line286*
+
 4.
 Line 143 Clustal-Omega is usually used as a multiple sequence aligner, can the authors provide details on what method it uses for producing a tree and whether branch lengths were provided to PAML or estimated in PAML?
+
+	*Nice catch! It is ClustalW2-Phylogeny. I have changed that in the text, as well as indicate that branch lengths are estimated in this software. Line 168*
 
@@
 
-Minor comments:
+	*I would rather not, given this is outside of the current project.
+I actually did not even calculate individual expression values.
+I suppose I could if you think it would add significantly to the manuscript (I don't think it does).*
 
-Line 95, do the authors mean PHRED < 2 or PHRED <20?
+>Minor comments:
 
-Line 109 should this say "using default settings"?
+6. Line 95, do the authors mean PHRED < 2 or PHRED <20?
 
-Validity of the findings
+	*Phred <2*
 
-Major comments:
+7. Line 109 should this say "using default settings"?
+
+	*Fixed*
+
+>Validity of the findings
+
+>Major comments:
 

### 2014-09-28 Pero_Transcriptome.tex
cont

@@
 
-To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted \textit{Dipodomys} rodents \citep{Marra:2014de}.
+To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted Heteromyid rodents \citep{Marra:2014de}.
 The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b \citep{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \citep{Aubry:2014en}).
@@
 To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above using alignments produced in MACSE.
-These alignments were free from indels and internal stop codons.
+These alignments were manually inspected, and were relatively free from indels and internal stop codons.
 We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.

### 2014-09-28 rebuttal.md
further

@@
 
-1. The authors mention calculating the site frequency spectrum for their data in line 133, did anything come of this analysis?
+8. The authors mention calculating the site frequency spectrum for their data in line 133, did anything come of this analysis?
 
-2.
+	*This SFS analysis was done as part of the calculation of Tajima's D. Nothing per se was done with the SFS itself.*
+
+9.
 Is anything known about the demographic history of this population?
@@
 If this data does not exist the authors should state that this and include demography as a possible explanation of their data in their conclusions about tables 3 and 4.
+
+	*Nothing specific is known of these populations, other than what can be inferred from paleo-climate models of desert expansion over the last 10k years.
+I have added a claus in the interpretation of Tajima's D to reflect the possibility that demography may impact these metrics.
+Line 257*
 
@@
 Line 200: Do the authors mean complete coding sequence or open reading frame rather than 'complete exons'?
-
-Comments for the author
-
-In this study the authors characterized the transcriptome of four separate tissues for an individual Peromyscus eremicus and conducted RNA-seq of kidney tissue of an additional 15 individuals to provide transcriptomic resources for the study of osmoregulation in this desert rodent.
-The study is sound in its methodology and provides a significant resource, however it does require some modifications and/or clarifications about some of the methodology and data before it could be published.
-In the above sections I provide suggestions and/or questions that should be addressed.
-
-© 2014, PeerJ, Inc. PO Box 614 Corte Madera, CA 94976, USA

### 2014-09-28 Pero_Transcriptome.tex
further

@@
 In addition to this, we estimated mean TPM (number of transcripts per million) for all transcripts.
-{\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TMP was the highest.
+{\hypertarget{Table 2}{Table 2}} consists of the 10 genes whose mean TPM was the highest.
 Several genes in this list are predominately  present in a single tissue type.
@@
 The distribution of the estimates of Tajima's D for all of the assembled transcripts is shown in Figure 2.
-The distribution is skewed toward negative values (mean=-0.89, variance=0.58), which is likely the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \citep{Chamary:2006db}.
+Although Tajima's D is known to be sensitive to demographic history, which is largely unknown for this population, the estimates may also be drive by patterns of selection.
+In general, the distribution is skewed toward negative values (mean=-0.89, variance=0.58), which may be the result of purifying selection, a model of evolution commonly invoked for coding DNA sequences \citep{Chamary:2006db}.
 \hyperlink{Table 3}{Table 3} presents the 10 transcripts whose estimate of Tajima's D is the greatest, while \hyperlink{Table 4}{Table 4} presents the 10 transcripts whose estimate of Tajima's D is the least.
-The former list of genes is likely to contain transcripts experiencing balancing selection in the studied population.
+The former list of genes may contain transcripts experiencing balancing selection in the studied population.
 This list includes, interestingly, genes obviously related to solute and water balance (e.g.

### 2014-09-28 Pero_Transcriptome.tex
maybe almost finished with revisions

@@
 To characterize natural selection on several genes related to water and ion homeostasis, we identified several of the transcripts identified as experiencing positive selection in a recent work on desert-adapted Heteromyid rodents \citep{Marra:2014de}.
-The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9) and the Vitamin D3 receptor (Vdr), were extracted from the dataset, aligned using the software MACSE version 1.01b \citep{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \citep{Aubry:2014en}).
+The coding sequences corresponding to these genes, Solute Carrier family 2 member 9  (Slc2a9), the Vitamin D3 receptor (Vdr) and several of the Aquaporin genes (Aqp1,2,4,9), were extracted from the dataset, aligned using the software MACSE version 1.01b \citep{Ranwez:2011kj} to homologous sequences in \textit{Mus musculus}, \textit{Rattus norvegicus}, \textit{Peromyscus maniculatus}, and \textit{Homo sapiens} as identified by the conditional reciprocal best blast procedure (CRBB, \citep{Aubry:2014en}).
 An unrooted gene tree with branch lengths was constructed using the online resource ClustalW2-Phylogeny (\url{http://www.ebi.ac.uk/Tools/phylogeny/clustalw2_phylogeny/}), and the tree and alignment were analyzed using the branch-site model (model=2, nsSites=2, fix\_omega=0 versus model=2, nsSites=2, fix\_omega=1, omega=1)  implemented in PAML version 4.8 \citep{Yang:2011bm,Yang:2007ki}.
@@
 From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be made available on Dryad).
-Of these 71626 sequences, 38221 were complete exons (containing both start and stop codons), while the others were either truncated at the 5-prime end (20239 sequences), the 3-prime end (6445 sequences), or were internal (6721 sequencing with neither stop nor start codon).
+Of these 71626 sequences, 38221 contained complete open reading frames (containing both start and stop codons), while the others were either truncated at the 5-prime end (20239 sequences), the 3-prime end (6445 sequences), or were internal (6721 sequencing with neither stop nor start codon).
 The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}..
@@
 
-To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we implemented the branch-site test as described above using alignments produced in MACSE.
+To begin to test the hypothesis that selection on transcripts related to osmoregulation is enhanced in the desert adapted \peer, we calculated Tajima's D as described above, and implemented the branch-site test using alignments produced in MACSE.
 These alignments were manually inspected, and were relatively free from indels and internal stop codons.
-We set the sequence corresponding to \peer\:for both Slc2a9 and Vdr as the foreground lineages in 2 distinct program executions.
-These two transcripts were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.
-The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0), indicating enhanced selection in \peer\: relative to the other lineages.
-The branch site test for positive selection conducted on the Vdr gene was non-significant (2$\Delta$Lnl=0.68, df=1, p=1).
-This limited analysis of selection is to be followed up by an analysis of genome wide patterns of natural selection.
+We set the sequence corresponding to \peer\:for both Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in 2 distinct program executions.
+These transcripts Slc2a9 and Vdr were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.
+The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0, Table 5), indicating enhanced selection in \peer\: relative to the other lineages.
+The branch site test for positive selection conducted on the Vdr and Aquaporin genes were non-significant.
+While the branch site test of positive selection is largely non-significant, estimating Tajima's D for these few candidate loci demonstrates that either a selective or demographic process may be influencing the genome at these functionally relevant sites.
 \\
+
+\textbf{\hypertarget{Table 5}{Table 5}} \\
+\begin{center}
+\begin{tabular}{ l c c c }
+\textbf{Transcript ID} & \textbf{Description} & \textbf{Tajima's D} & \textbf{Branch Site Test p.value} \\
+\hline
+Transcript\_106085 & Slc2a9 & 2.15 & p=0\\
+Transcript\_114624 & Vdr & 1.97 & p=1 \\
+Transcript\_128972 & Aqp1 & 1.39 & p=1 \\
+Transcript\_33960 & Aqp2 & 1.78 & p=1 \\
+Transcript\_22154 & Aqp4 & 2.10 & p=1 \\
+Transcript\_107677 & Aqp9 & 2.06 & p=1 \\
+
+ \end{tabular}
+\begin{quote}
+\small{Table 5. Several candidate genes were evaluated using Tajima's D and the branch site test implemented in PAML. }
+\end{quote}  
+\end{center}
 
@@
 
+This manuscript was greatly improved by careful review from C.
+Titus Brown, Elijah Lowe, and an anonymous reviewer, as well as by Matthew Hahn, who provided feedback on an earlier version of the manuscript posted on bioRxiv.
+
 %MDM would like to acknowledge the support of his wife and children.

### 2014-09-28 rebuttal.md
maybe almost finished with revisions
done with revisions, still need to check refs.

@@
 	*I did not conduct an assembly of the pooled samples is because I believe that tissue-specific isoforms may be reconstructed with more fidelity in this manner.
-I make a statement to this effect on line 103:105*
+I make a statement to this effect on line 111.
+I do trade off the potential for reconstructing additional low-coverage transcripts, but for the sake of future studies, more accurate isoform reconstruction is more important.
+Therefore, this is a tradeoff I am willing to take.*
 
 	*I did not use the 15 replicate individuals in the assembly for concerns that the added polymorphism would increase run time, hardware requirements, and would decrease assembly contiguity - all effects related to a more complex de Bruijn graph.
-Line 99:100*
+Line 114*
 
@@
 I do wonder why this organ produced a larger number of contigs in the assembly, but a similar number of transcripts after filtering.
-I admit that I do not have a good explanation, and I hope a reader will..*
+I admit that I do not have a good explanation, and I hope a reader will..
+The data was not of lower quality, nor are any of the assembly metrics.*
 
@@
 I think it is likely that I could have recovered more contigs using Pfam, but I would worry that the sensitivity of a HMM based search may result in the recovery of 'false' transcripts, which I am relatively intolerant of.
-I am electing to not filter based on Pfam, understanding the implications of this.*
+I am electing to not filter based on Pfam, understanding the implications and potential biases of employing this strategy, though I believe them to be minimal, given that nearly all (>92%) raw reads map back to the assembly.*
 
@@
 	*I have added a few more genes, Aqp 1,2,4,9.
-There is not evidence of positive selection in the eremicus lineage for any of these genes, as expected.
+(New table 5).
+I did not elect to run the PAML branch site analysis on the genes identified in the analysis of Tajima's D, because for these genes, there is no a priori expectation that selection on these genes is acting differently in eremicus that in the other lineages, as there might be for genes related to osmoregulation.
+For the tested Aquaporins, there is no evidence of positive selection in the eremicus lineage for any of these genes, as expected.
 The de novo (e.g.
 NOT focused on candidate genes) exploration of positive selection on P.
-eremicus will be presented in a later work, in connection with a genome sequence*
+eremicus will be presented in a later work, in connection with a newly produced genome sequence.*
 
@@
 
-Overall I believe it is a good paper with interesting analysis, and cool results.
+---
 
@@
 After all, these are animals that live in the desert.
-Rain may happen on rare occasions, but by definition deserts are habitats exceptionally devoid of naturally-occuring drinking water.*
+Rain may happen on very rare occasions, but by definition deserts are habitats exceptionally devoid of naturally-occuring drinking water.*
 
@@
 For instance, I specify that the 4 reference tissue samples were sequenced on 2 lanes of a Hiseq 2500.
-The 15 replicate samples were sequence across several lanes, so this number (how many individuals sequenced per lane) is not representative.
-Instead, the number of reads seems like the more informative number.
-*
+The 15 replicate samples were sequence across several lanes, sometimes with other samples unrelated to this project, so this number (how many individuals sequenced per lane) is not representative.
+Instead, the number of reads seems like the more informative number.*
 
@@
 
-	*I have added the definition at the 1st use of the abbreviation.*
+	*I have added the definition at the 1st use of the abbreviation, on line 125.*
 
@@
 
-	*Tricky grammer issue. I have changed these terms to their plural*
+	*Tricky grammar issue. I have changed these terms to their plural*
 
@@
 
-	*Genes were given symbols when symbols exist. For some, e.g. h-2 class I histocompatibility antigen XM_006997718.1 has the symbol "LOC102911283". This does not seem informative, so it was left off.*
+	*Genes were given symbols when symbols exist.
+For a few, e.g.
+h-2 class I histocompatibility antigen XM_006997718.1 has the symbol "LOC102911283".
+This does not seem informative, so it was left off.
+Does this make sense to do so?*
 
@@
 
-	*Excellent point. I have added the citation here and changed Dipodomys to Hetermyid to more accurately reflect the work done in that paper.*
+	*Excellent point. I have added the citation here and changed Dipodomys to Heteromyid to more accurately reflect the work done in that paper.*
 
@@
 	*The alignments were visually inspected.
-While some indels do occur (especially in Homo relative to the other rodents), these all occur in units of 3, which to me at least suggests that they are likely accurate.
-The alignment column in which premature stop codons are treated as missing data in PAML.
+While some indels do occur (especially in Homo relative to the other rodents), these all occur in units of 3nt, which to me at least suggests that they are likely accurate.
+The alignment column in which premature stop codons are treated as missing data in PAML, though this was not common in these analyses.
 Line286*
@@
 
-	*Nice catch! It is ClustalW2-Phylogeny. I have changed that in the text, as well as indicate that branch lengths are estimated in this software. Line 168*
+	*Nice catch! It is ClustalW2-Phylogeny. I have changed that in the text, added a link to that webpage, as well as indicate that branch lengths are estimated in this software. Line 168*
 
@@
 
-	*This SFS analysis was done as part of the calculation of Tajima's D. Nothing per se was done with the SFS itself.*
+	*This SFS analysis was done as part of the calculation of Tajima's D. Nothing per se was done with the SFS itself. I will add this to Dryad, in case somebody finds it interesting or useful.*
 
@@
 	*Nothing specific is known of these populations, other than what can be inferred from paleo-climate models of desert expansion over the last 10k years.
-I have added a claus in the interpretation of Tajima's D to reflect the possibility that demography may impact these metrics.
+I have added a statement in the interpretation of Tajima's D to reflect the possibility that demography may impact these metrics.
 Line 257*
@@
 
-Minor comments:
+	*Excellent idea. See the new table 5*
 
-Line 200: Do the authors mean complete coding sequence or open reading frame rather than 'complete exons'?
+> Minor comments:
+
+4. Line 200: Do the authors mean complete coding sequence or open reading frame rather than 'complete exons'?
+
+	*Yes, the wording has been changed to 'open reading frame'*

### 2014-09-28 Pero_Transcriptome.tex
done with revisions, still need to check refs.

@@
 submitted for Illumina sequencing using two lanes of 150nt paired end sequencing employing the rapid-mode of the HiSeq 2500 sequencer at The Hubbard Center for Genome Sciences (University of New Hampshire).
-The remaining 15 libraries were similarly multiplexed and sequenced in a mixture of 100nt paired and single end sequencing runs across several lanes of an Illumina HiSeq 2000 at the Vincent G.
+The remaining 15 libraries were multiplexed and sequenced in a mixture of 100nt paired and single end sequencing runs across several lanes of an Illumina HiSeq 2000 at the Vincent G.
 Coates Genome Center (University of California, Berkeley).
@@
 Libraries were constructed as per the standard Illumina protocol and sequenced as described above.
-The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available on the Short Read Archive accession XXX).
+The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available as part BioProject PRJNA242486).
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of the total number of reads.

### 2014-09-28 Pero_Transcriptome.tex
finishing up refs
final print
ref changing

@@
 \usepackage{lineno}
+\usepackage{url}
 
@@
 Matthew D. MacManes$^{1}$, 
-Michael B. Eisen $^{2}$, 
+Michael B. Eisen $^{2}$
 \\
@@
 The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \citep{MacManes:2014io} and Mbandi \citep{Christoffels:2014bg}.
-Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \citep{Lohse:2012fg}.
+Specifically, adapter sequence contamination and low quality nucleotides (defined as \textsc{Phred} \textless 2) were removed using the program Trimmomatic version 0.32  \citep{Bolger:2014ek}.
 Reads from each tissue were assembled using the Trinity version released 17 July 2014 \citep{Haas:2013jq}.
 We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt.
-We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy and MacManes unpublished observations}.
+We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy}.
 Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.
@@
 % The bibtex filename
-\bibliography{/Users/macmanes/Documents/pero_transcriptome/biblio.bib}
+\bibliography{/Users/macmanes/Documents/pero_transcriptome/biblio2.bib}
 

### 2014-09-28 Pero_Transcriptome.tex
restructure repo for publication

@@
 These alignments were manually inspected, and were relatively free from indels and internal stop codons.
-We set the sequence corresponding to \peer\:for both Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in 2 distinct program executions.
+We set the sequence corresponding to \peer\:for both Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in six distinct program executions.
 These transcripts Slc2a9 and Vdr were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.

### 2014-09-29 Pero_Transcriptome.tex
update

@@
 
-The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#error-correction}.
 The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \citep{MacManes:2014io} and Mbandi \citep{Christoffels:2014bg}.
@@
 Reads from each tissue were assembled using the Trinity version released 17 July 2014 \citep{Haas:2013jq}.
-We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt.
+We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#trinity-assemblies}.
 We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy}.
@@
 We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
-eremicus} dataset likely to be biological in origin.
+eremicus} dataset likely to be biological in origin \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#blasting}.
 This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est-}.
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \citep{Mi:2004iv}.
-Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko}.
-Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/})\\
+Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko} \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#hmmer3pfam}.
+Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/}) \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#transdecoder-}\\
 
@@
 Sequence data were mapped to the reference transcriptome using bwa-mem.
-The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz}.
+The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz} using instructions found at \url{http://popgen.dk/angsd/index.php/Tajima}.
 \\

### 2014-09-29 Pero_Transcriptome.tex
AM

@@
 
-The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#error-correction}.
+The raw sequence reads corresponding to the four tissue types were error corrected using the software bless version 0.17 \citep{Heo:2014cb} using kmer=25, based on the developer's default recommendations (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#error-correction}).
 The error-corrected sequence reads were adapter and quality trimmed following recommendations from MacManes \citep{MacManes:2014io} and Mbandi \citep{Christoffels:2014bg}.
@@
 Reads from each tissue were assembled using the Trinity version released 17 July 2014 \citep{Haas:2013jq}.
-We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#trinity-assemblies}.
+We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#trinity-assemblies}).
 We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy}.
@@
 This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est-}.
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est-}).
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 Among other things, the Ensemble transcript identifiers were used in the analysis of gene ontology conducted in the PANTHER package \citep{Mi:2004iv}.
-Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko} \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#hmmer3pfam}.
-Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/}) \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#transdecoder-}\\
+Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko} (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#hmmer3pfam}).
+Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/}) (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#transdecoder-})\\
 

### 2014-09-29 Pero_Transcriptome.tex
fix names, again..
for upload to PeerJ

@@
 We used a blastN (version 2.2.29+) procedure (default settings, evalue set to 10$^{-10}$) to identify contigs in the \textit{P.
-eremicus} dataset likely to be biological in origin \url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#blasting}.
+eremicus} dataset likely to be biological in origin (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#blasting}).
 This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
-We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est-}).
+We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est}).
 This multi-fasta file was used for all subsequent analyses, including annotation and mapping.
@@
 Next, because rapidly evolving nucleotide sequences may evade detection by blast algorithms, we used HMMER3 version 3.1b1 \citep{Wheeler:2013gj} to search for conserved protein domains contained in the dataset using the Pfam database \citep{Punta:2012ko} (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#hmmer3pfam}).
-Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/}) (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#transdecoder-})\\
+Lastly, we extracted putative coding sequences using Transdecoder version 4Jul2014 (\url{http://transdecoder.sourceforge.net/}) (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#transdecoder})\\
 
@@
 Sequence data were mapped to the reference transcriptome using bwa-mem.
-The alignments were sorted and converted to BAM format, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz} using instructions found at \url{http://popgen.dk/angsd/index.php/Tajima}.
+The alignments were sorted and converted to BAM format using the samtools software package \citep{Li:2009ka}, then passed to the program \textsc{Angsd} version 0.610, which was used for calculating the folded site frequency spectrum (SFS) and Tajima's D \citep{Korneliussen:2013uz} using instructions found at \url{http://popgen.dk/angsd/index.php/Tajima}.
 \\
@@
 \textsc{Peer360 Testes} & 32M PE/SS  & SRR1575398 \\
-\textsc{Peer360 Liver} & 53M PE /SS & SRR1575397 \\
+\textsc{Peer360 Liver} & 53M PE/SS & SRR1575397 \\
 \textsc{Peer360 Kidney} & 56M PE/SS  & SRR1575396 \\
@@
 
-After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 123,123 putative transcripts remained (to be made available on Genbank, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}.).
+After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 122,584 putative transcripts remained (to be made available on Genbank, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}.).
 From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be made available on Dryad).
@@
 The Tajima's D statistic was calculated for all  transcripts covered by at least 14 of the 15 individuals.
-In brief, a negative Tajima's D - a result of lower than expected average heterozygosity - is often associated with purifying or directional selection, recent selective sweep or population bottleneck.
+In brief, a negative Tajima's D - a result of lower than expected average heterozygosity - is often associated with purifying or directional selection, recent selective sweep or recent population expansion, or a complex combination of these forces.
 In contrast, a positive value for Tajima's D represents higher than expected average heterozygosity, often associated with balancing selection.
@@
 These alignments were manually inspected, and were relatively free from indels and internal stop codons.
-We set the sequence corresponding to \peer\:for both Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in six distinct program executions.
+We set the sequence corresponding to \peer\:for Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in six distinct program executions.
 These transcripts Slc2a9 and Vdr were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.

### 2014-10-08 Pero_Transcriptome.tex
accepted!

@@
 We used flags to indicate the stranded nature of sequencing reads and set the maximum allowable physical distance between read pairs to 999nt (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#trinity-assemblies}).
-We elected to assembly reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy}.
-Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity that if all tissues were assembled together.
+We elected to assemble reads derived from a single deeply sequenced individual (Peer360, a male) to reduce polymorphism and thus the complexity of the de Bruijn graph, which has important implications for runtime, hardware requirements \citep{Lowe:2014iv,Pop:2009dp}, and assembly contiguity \citep{Vijay:2012gy}.
+Individual tissues were assembled independently, as we hypothesize that tissue specific isoforms would be reconstructed with higher fidelity than if all tissues were assembled together.
 
@@
 eremicus} dataset likely to be biological in origin (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#blasting}).
-This procedure, when a reference dataset is available, retains more putative transcripts that a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
+This procedure, when a reference dataset is available, retains more putative transcripts than a strategy employing expression-based filtering (remove if transcripts per million (TPM) \textless 1  \citep{MacManes:2012bu}) of the raw assembly.
 We then concatenated the filtered assemblies from each tissue into a single file and reduced redundancy using the software cd-hit-est version 4.6 \citep{Li:2006hr} using default settings, except that sequences were clustered based on 95\% sequence similarity (\url{https://github.com/macmanes/pero_transcriptome/blob/master/analyses.md#cd-hit-est}).
@@
 Libraries were constructed as per the standard Illumina protocol and sequenced as described above.
-The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available as part BioProject PRJNA242486).
+The number of reads per library varied from 56 million strand-specific paired-end reads in Peer360 kidney, to 9 million single-end reads in Peer321 (\hypertarget{Table 1}{Table 1}, available as part of BioProject PRJNA242486).
 Adapter sequence contamination and low-quality nucleotides were eliminated, which resulted in a loss of \textless 2\% of the total number of reads.
@@
 
-Transcriptome assemblies for each tissue type was accomplished using the program Trinity \citep{Haas:2013jq}.
+Transcriptome assemblies for each tissue type were accomplished using the program Trinity \citep{Haas:2013jq}.
 The raw assemblies for brain, liver, testes, and kidney contained 185425, 222096, 180233, and 514091 assembled sequences respectively.
@@
 Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous.
-66324 transcripts are expressed on all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
+66324 transcripts are expressed in all tissue types, while 13086 are uniquely expressed in the kidney, 2222 in the testes, 3580 in the brain, and 2798 in the liver.
 The kidney appears to an outlier in the number of unique sequences, though this could be the result of the recovery of more lowly expressed transcripts or isoforms.
@@
 We set the sequence corresponding to \peer\:for Slc2a9, Vdr, and several of the Aquaporin genes (Aqp1,2,4,9) as the foreground lineages in six distinct program executions.
-These transcripts Slc2a9 and Vdr were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.
+The transcripts Slc2a9 and Vdr were chosen specifically because they - the former significantly - were recently linked to osmoregulation in a desert rodent \citep{Marra:2014de}.
 The test for Slc2a9 was highly significant (2$\Delta$Lnl=51.4, df=1, p=0, Table 5), indicating enhanced selection in \peer\: relative to the other lineages.
-The branch site test for positive selection conducted on the Vdr and Aquaporin genes were non-significant.
+The branch site tests for positive selection conducted on the Vdr and Aquaporin genes were non-significant.
 While the branch site test of positive selection is largely non-significant, estimating Tajima's D for these few candidate loci demonstrates that either a selective or demographic process may be influencing the genome at these functionally relevant sites.

### 2014-10-09 Pero_Transcriptome.tex
adding Dryad links

@@
 Together, these statistics suggest that the tissue-specific transcriptomes are of extremely high quality.
-All tissue-specific assemblies are to be made available on Dryad, and until then are stored on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}).
+All tissue-specific assemblies are available on Dryad (\url{http://dx.doi.org/10.5061/dryad.qf1dp}).
 \\
@@
 
-We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues (Pero.tissue.xprs, will be made available on Dryad, until then on Dropbox (\url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0})).
+We then estimated gene expression on each of these tissue-specific datasets, which allowed us to understand expression patterns in the multiple tissues (Pero.tissue.xprs, available on Dryad (\url{http://dx.doi.org/10.5061/dryad.qf1dp})).
 Specifically, we constructed a Venn diagram ({\hypertarget{Figure 1}{Figure 1}}) that allowed us to visualize the proportion of genes whose expression was limited to a single tissue and those whose expression was ubiquitous.
@@
 
-After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 122,584 putative transcripts remained (to be made available on Genbank, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}.).
-From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, to be made available on Dryad).
+After expression estimation, the filtered assemblies were concatenated together, and after the removal of redundancy with cd-hit-est, 122,584 putative transcripts remained (Pero.BLTK.fasta - available on Dryad: \url{http://dx.doi.org/10.5061/dryad.qf1dp}.).
+From this filtered concatenated dataset, we extracted 71626 putative coding sequences (72Mb, available on Dryad).
 Of these 71626 sequences, 38221 contained complete open reading frames (containing both start and stop codons), while the others were either truncated at the 5-prime end (20239 sequences), the 3-prime end (6445 sequences), or were internal (6721 sequencing with neither stop nor start codon).
-The results of a Pfam search conducted on the predicted amino acid sequences will be found on Dryad, and until then are stored on Dropbox \url{https://www.dropbox.com/sh/2jwzd8p6n6eluco/AAAO3nSdXb_u4wtQZRTwqW9ia?dl=0}..
+The results of a Pfam search conducted on the predicted amino acid sequences are on Dryad (\url{http://dx.doi.org/10.5061/dryad.qf1dp}).
 \\
